\section{Introduction}
Modern code-generation benchmarks~\cite{chen2021codex,austin2021program} show that large language models (LLMs) can solve standard programming tasks with high accuracy.
This creates a critical challenge for programming education: if LLMs solve assignments faster and more accurately than students, how can instructors meaningfully assess student understanding?

This paper studies a complementary but understudied question:
\begin{quote}
\small\textit{Is it possible to design a programming language that is human-interpretable but systematically defeats LLM next-token-prediction priors---making it resistant to AI-based code generation?}
\end{quote}

\noindent\textbf{The core asymmetry.}
The key insight is that humans and LLMs fail in fundamentally different ways when given an unfamiliar language rule.
A human student who reads the rule \emph{``if-blocks execute when the condition is \textbf{false}''} can immediately apply it---the rule is explicit, verifiable, and overrides prior habit.
An LLM given the identical rule in its prompt instead produces standard Python: its next-token prediction (NTP) prior, built from billions of Python training tokens, systematically overrides the stated instruction.
This asymmetry---\emph{humans follow novel rules; LLMs revert to training priors}---is the foundation of our resistance framework and requires no additional experiment to accept: it follows directly from the mechanics of NTP-based generation.

We decompose this into three sub-questions:
\begin{enumerate}
  \item Does changing token identifiers (L1: alias substitution) create effective LLM resistance?
  \item Does inverting syntactic structure (L2) create effective resistance?
  \item Does inverting semantic behavior (L3/L4) create deeper resistance?
\end{enumerate}

Our key empirical findings are:
\begin{itemize}
  \item \textbf{L1 (token substitution):} Partially effective---keyword leakage rate (KLR) ranges from 0.26 to 0.69 across models. Models revert to Python priors at measurable rates.
  \item \textbf{L2 (syntax inversion):} Not effective---models adopt new syntax completely (SIR=0.0) when given explicit rules.
  \item \textbf{L3 (semantic inversion, explicit):} Capacity-dependent---weaker models fail completely (PPR=1.0 for \texttt{gpt-4o-mini}), while stronger models succeed (PPR=0.0 for \texttt{gpt-4o} and \texttt{gpt-4.1-mini}).
  \item \textbf{L4 (semantic inversion, implicit):} Effective against all tested models on prior-entrenched tasks (e.g., Fibonacci)---even strong models exhibit \emph{pattern blindness}. However, capable models (gpt-5.4, o3) succeed on complex algorithmic tasks (merge sort, BFS) where prior entrenchment is shallow.
\end{itemize}
A key boundary condition: L4 resistance requires tasks with deep, uniform Python priors. We identify \emph{prior entrenchment depth} as a third design axis beyond perturbation level and rule delivery mode.

We also identify three distinct LLM failure modes: \emph{prior dominance}, where the prior overrides stated rules; \emph{pattern blindness}, where the model fails to extract semantic rules from examples; and \emph{operational substitution}, where the model achieves correct output via arithmetic workaround rather than adopting the semantic rule.\footnote{Operational substitution was observed on \texttt{count\_vowels} for \texttt{gpt-4o} at $n=50$.}

This framework provides both a theoretical taxonomy for LLM-resistant language design and practical tools for constructing programming assessments that remain meaningful in the LLM era.
