\section{Related Work}

\subsection{Code Generation Benchmarks and Evaluation}
HumanEval~\cite{chen2021codex} and MBPP~\cite{austin2021program} established functional correctness as the standard metric for code-generation evaluation. More recent benchmarks scale and diversify tasks: BigCodeBench~\cite{bigcodebench2024} adds complex, multi-function instructions, and LiveCodeBench~\cite{livecodebench2024} avoids contamination through continuous updating. Across all these benchmarks, Python dominates training corpora, embedding strong token-level priors into all models evaluated. This motivates our L1 baseline: even with explicit alias instructions, models leak Python keywords at 26--69\%.

\subsection{Code-Oriented LLM Architectures and Training}
Pre-trained code models---CodeBERT~\cite{feng2020codebert}, CodeT5~\cite{wang2021codet5}, and InCoder~\cite{fried2022incoder}---demonstrate that NTP objectives on Python code reinforce keyword position priors. More recent instruction-tuned models (WizardCoder~\cite{luo2023wizardcoder}, DeepSeek-Coder~\cite{guo2024deepseek}) amplify these priors through even larger code corpora. This trajectory predicts our finding: \emph{stronger code LLMs do not necessarily mean better alias compliance}---gpt-5.4-mini (KLR=0.67) underperforms gpt-4o (KLR=0.37).

\subsection{Instruction-Following in Code (2024--2025)}
CodeIF~\cite{codeif2025} and MultiCodeIF~\cite{multicode2025} introduce benchmarks measuring LLM adherence to developer-specified constraints in code generation tasks.
Critically, MultiCodeIF reports that models perform relatively well on \emph{explicit, concrete} constraints but degrade substantially on \emph{implicit or abstract} constraints, and that performance drops sharply as the number of hierarchical constraints increases.
This explicit--implicit asymmetry is the closest mechanistic evidence in the literature to our L3/L4 findings: models can follow explicit semantic inversion rules (L3, with some model-capacity dependence), yet fail entirely when the same semantic inversion is delivered only through examples without explicit articulation (L4, pattern blindness across all tested models).
The key distinction is scope: CodeIF and MultiCodeIF benchmark \emph{constraint-following} within standard programming tasks, whereas our work \emph{designs} a programming language whose semantic structure systematically conflicts with NTP priors, making resistance a structural property rather than a task-level benchmark artifact.

\subsection{Instruction Hierarchy and Prior Conflicts}
Chen et al.~\cite{chen2024instructhierarchy} propose an instruction hierarchy to resolve conflicts between system-level and user-level instructions. Their framework addresses natural-language directive conflicts; ours involves programming-language semantic rule conflicts---a dimension not studied in alignment work.
Xie et al.~\cite{xie2024conflict} and Wu et al.~\cite{wu2024contextparametric} show that when context conflicts with parametric knowledge, LLMs often default to pretraining priors. Zhang et al.~\cite{zhang2024conflict} survey this knowledge conflict problem broadly. Shi et al.~\cite{contextinvert2024} demonstrate LLMs are easily distracted by irrelevant context; our L4 condition represents the inverse: models \emph{fail to use} relevant semantic cues provided in examples.

\subsection{NTP Priors and In-Context Learning}
Wang et al.~\cite{wang2024ntp} formalize a ``law of next-token prediction'' providing theoretical grounding for how NTP training shapes output distributions. Min et al.~\cite{min2022rethinking} show ICL demonstrations affect outputs primarily through distributional surface patterns rather than semantic content---explaining why models fail to extract inverted semantics from L4 examples. Dong et al.~\cite{dong2022survey} survey ICL comprehensively. Lampinen et al.~\cite{anotb2024} document ``A-Not-B'' errors where LLMs fail to suppress previously established response patterns under context changes, directly analogous to our L4 pattern blindness.

\subsection{Counterfactual and Semantic Reasoning Failures}
Zhao et al.~\cite{counterfact2024} show LLMs struggle when counterfactual reasoning requires overriding parametric knowledge---the same mechanism we observe at L3 for weak models. Turpin et al.~\cite{semanticinvar2024} show LLM self-explanations are often unfaithful to actual behavior. Valmeekam et al.~\cite{reasoningrigidity2024} document reasoning rigidity---models revert to memorized strategies under redefined rules---providing further precedent for our pattern blindness finding.

\subsection{Adversarial Perturbation and Code Resistance}
PromptBench~\cite{zhu2023promptrobust} demonstrates LLM fragility under text perturbations. Pan et al.~\cite{pan2024adversarial} show semantically aligned code prompt perturbations cause up to 42\% accuracy drops. Jha et al.~\cite{coderobbery2024} use structured adversarial prompts to break LLM reasoning in code tasks. Khan et al.~\cite{khan2024cheatingllm} specifically design prompt perturbations to reduce LLM correctness on programming assignments by 77\%, most directly related to our goal. Our approach is complementary: instead of perturbing prompts, we design a confusion language whose NTP prior conflict is a \emph{structural property}, making it principled and reproducible.

\subsection{DSL and Constrained Code Generation}
Wu et al.~\cite{wu2024lowresource} survey LLM performance on low-resource and domain-specific languages, confirming that Python-trained models underperform on less common languages. ConCodeEval~\cite{concode2024} benchmarks code constraint comprehension in DSLs; L2CEval~\cite{l2ceval2024} evaluates language-to-code generation across specialized languages. Constrained decoding~\cite{dong2024xgrammar,muendler2025type} enforces grammar-level output validity. Our L2 experiment shows models adopt explicit syntax rules entirely, but semantic constraints (L3/L4) prove far more resistant---showing grammar-level constraints are insufficient without semantic-level challenges.

\subsection{Real-World Engineering Benchmarks}
SWE-bench~\cite{jimenez2024swebench} reveals that LLMs fail to generalize from canonical benchmarks to real repository-level tasks---confirming that high performance on standard benchmarks does not indicate robustness to structural novelty, as we observe in our L1~L4 evaluations.

\subsection{LLM in Programming Education}
Dakhel et al.~\cite{dakhel2023github} show LLM coding tools are an ``asset or liability'' depending on student expertise. Lau and Guo~\cite{lau2023learners} document novice programmers adopting LLM outputs uncritically. Prather et al.~\cite{prather2023robots} show mixed student perceptions. Liffiton et al.~\cite{liffiton2023codehelp} propose guardrailed LLM tutoring. These studies confirm a shared concern: LLM integration can undermine independent skill development.

\subsection{Academic Integrity and LLM-Resistant Assessment}
Novak and McDanel~\cite{novak2025llmhw} propose contextual, personalized assignments to resist LLM cheating---the most directly related paper to our educational goal---but without a formal theory of what features drive LLM failure. Code perplexity detection~\cite{codeplagiarism2024} is brittle; our taxonomy offers a principled design-based solution. RLHF-based alignment~\cite{ouyang2022instruct} improves instruction-following but, as our L3/L4 results show, remains insufficient when semantic inversions must be extracted from examples.

\subsection{Specification Gaming and Reward Hacking}
Krakovna et al.~\cite{krakovna2020specgaming} catalogue specification gaming behaviors where agents exploit loopholes rather than following intended objectives. Our ``operational substitution'' failure mode---where models produce correct output via arithmetic workaround instead of adopting the semantic rule---is a code-generation analogue of specification gaming: the model satisfies the output specification without acquiring the intended semantic competence.

\subsection{Adversarial Robustness in Code Generation}
Prior adversarial code work (PromptBench~\cite{zhu2023promptrobust}, Pan et al.~\cite{pan2024adversarial}) focuses on \emph{syntactic} perturbations---renaming variables, reordering tokens---that degrade generation quality. Our approach targets \emph{semantic} inversion: the surface syntax remains standard Python, but the conditional semantics are reversed. This is a fundamentally different threat model---syntactic perturbation attacks the input representation, while semantic inversion attacks the model's prior over program behavior.

\subsection{Connection to Prompt Injection and Jailbreaking}
L3/L4 share structural similarity with prompt injection~\cite{chen2024instructhierarchy}: both present in-context instructions that conflict with the model's default behavior. However, the goal is inverted---in jailbreaking, the attacker wants the model to \emph{override} its safety prior; in our framework, we want the NTP prior to \emph{win}, producing standard Python despite the inverted-semantics instruction. L4 pattern blindness thus provides evidence that prior override is asymmetric: safety alignment can be bypassed by sufficiently crafted prompts, but code-generation priors resist override even when the user \emph{wants} compliance.

\subsection{Research Gap}
Prior adversarial work degrades generation quality without principled language design. Prior DSL work measures performance without explaining failure mechanisms. Prior knowledge-conflict work focuses on factual knowledge, not programming-language semantics. Our contribution closes these gaps: a theory-grounded L1--L4 taxonomy that characterizes LLM failure by mechanism (prior dominance vs.\ pattern blindness) and guides principled LLM-resistant language design.
