\section{Method}
\subsection{Research Questions and Hypotheses}
\textbf{RQ1}: At what language design level does LLM resistance become robust?
\textbf{RQ2}: What failure mechanism is active at each level---prior dominance or pattern blindness?

\textbf{H1}: Token-level alias substitution (L1) creates partial but not complete resistance.
\textbf{H2}: Syntactic inversion (L2) is insufficient---models adopt new syntax when given explicit rules.
\textbf{H3}: Semantic inversion (L3/L4) provides robust resistance, but differs in mechanism by model capacity.

\subsection{Language Level Taxonomy}
We design a four-level taxonomy along two axes: (1) \emph{perturbation depth} (token, syntax, semantic) and (2) \emph{rule delivery} (explicit vs.\ implicit). Table~\ref{tab:levels} summarizes each level.

\begin{table}[!t]
\centering
\caption{Four-level confusion language taxonomy. Each level targets a distinct linguistic dimension and delivery modality. L4 is the only level where all tested models fail regardless of capacity.}
\label{tab:levels}
\footnotesize
\begin{tabular}{@{}clp{1.5cm}p{2.8cm}@{}}
\toprule
\textbf{Level} & \textbf{Perturbation} & \textbf{Delivery} & \textbf{Observed failure mode} \\
\midrule
L1 & Token alias & Explicit & Prior dominance (partial) \\
L2 & Syntax inv. & Explicit & None---adopted \\
L3 & Semantic inv. & Explicit & Capacity-dependent \\
L4 & Semantic inv. & Implicit & Pattern blindness \\
\bottomrule
\end{tabular}
\par\smallskip
\raggedright\scriptsize\textit{Note:} L4 experiments use three distinct sub-conditions across sections: (1)~strict example-only ablation (§\ref{sec:l4_ablation}, $n=50$), the canonical L4 condition; (2)~partially-annotated multi-task (§\ref{sec:l4_multitask}), a weakened form with inline comment hints; and (3)~hard-task extension (§\ref{sec:l4_hard}), example-only on algorithmically complex tasks. Findings differ by sub-condition; see §\ref{sec:nonmonotonic} for synthesis.
\end{table}

Figure~\ref{fig:taxonomy} illustrates the two-axis design space and the location of each level.
The two gray cells (Token/Implicit and Syntax/Implicit) are theoretically possible configurations that we did not study; Token/Implicit would expect higher KLR than L1-Explicit (weaker contextual cues), while Syntax/Implicit would expect higher SIR. We leave these as open positions in the design space for future work.

\begin{figure}[!t]
\centering
\begin{tikzpicture}[x=2.2cm, y=1.4cm, font=\small,
  cell/.style={draw, rounded corners=3pt, minimum width=1.6cm,
    minimum height=0.9cm, align=center, text width=1.4cm,
    line width=0.4pt}]
  % Grid lines
  \draw[gray!30, line width=0.3pt] (0,0) grid (3,2);
  % Axis labels
  \node[below, font=\scriptsize\sffamily] at (0.5,-0.1) {Token};
  \node[below, font=\scriptsize\sffamily] at (1.5,-0.1) {Syntax};
  \node[below, font=\scriptsize\sffamily] at (2.5,-0.1) {Semantic};
  \node[left, font=\scriptsize\sffamily, align=right] at (-0.05,1.5) {Implicit};
  \node[left, font=\scriptsize\sffamily, align=right] at (-0.05,0.5) {Explicit};
  \node[below, font=\footnotesize\bfseries\sffamily] at (1.5,-0.55) {Perturbation Depth};
  \node[left, font=\footnotesize\bfseries\sffamily, rotate=90] at (-0.7,1) {Rule Delivery};
  % Cells — color code: green=resistant, yellow=partial, red=not resistant
  \node[cell, fill=yellow!25, draw=yellow!60!black]
    at (0.5,0.5) {\textbf{L1}\\{\scriptsize Prior dom.}};
  \node[cell, fill=red!12, draw=red!40!black]
    at (1.5,0.5) {\textbf{L2}\\{\scriptsize Adopted}};
  \node[cell, fill=yellow!25, draw=yellow!60!black]
    at (2.5,0.5) {\textbf{L3}\\{\scriptsize Cap.-dep.}};
  \node[cell, fill=green!25, draw=green!50!black]
    at (2.5,1.5) {\textbf{L4}\\{\scriptsize Patt.\ blind}};
  \node[cell, fill=gray!15, draw=gray!50, text=black!55,
    minimum width=1.8cm, minimum height=0.55cm, inner sep=3pt]
    at (0.5,1.5) {\scriptsize\itshape not studied};
  \node[cell, fill=gray!15, draw=gray!50, text=black!55,
    minimum width=1.8cm, minimum height=0.55cm, inner sep=3pt]
    at (1.5,1.5) {\scriptsize\itshape not studied};
\end{tikzpicture}
\caption{Two-axis confusion language design space. Perturbation depth (token/syntax/semantic) and rule delivery (explicit/implicit) together determine LLM failure mode. L4 (dark red) occupies the hardest cell: semantic perturbation without explicit rule---the only condition where all tested models fail.}
\label{fig:taxonomy}
\end{figure}

\subsection{L1: Token Alias Substitution}
Python keywords (\texttt{if}, \texttt{for}, \texttt{in}, \texttt{return}, \texttt{def}) are replaced with aliases. Constraints: one-to-one mapping, no alias collision, token-boundary replacement, literal/comment protection.

\subsection{L2: Syntax Structure Inversion}
Function definitions, conditionals, loops, and block delimiters are structurally inverted relative to Python. Example: \texttt{def fib(n):} becomes \texttt{:define fib [n] ->}. Model is given explicit grammar rules.

\subsection{L3: Semantic Inversion (Explicit Rules)}
Python surface syntax is retained but the semantics of if-blocks are inverted: the indented block executes when the condition is \emph{false}. Model is explicitly told this rule. Figure~\ref{fig:semantic_inversion} illustrates the concept.

\begin{figure}[!t]
\centering
\begin{tcolorbox}[colback=gray!8, colframe=black!60,
  title=\textbf{Python (standard)}, fonttitle=\small\bfseries, boxrule=0.5pt]
\begin{lstlisting}[language=Python,
  basicstyle=\ttfamily\footnotesize,
  columns=fixed,keepspaces=true,tabsize=4]
def fib(n):
    if n <= 1:
        return n
    return fib(n-1) + fib(n-2)
\end{lstlisting}
\tcblower
\small\textit{Standard semantics: if-block runs when condition is TRUE}
\end{tcolorbox}
\vspace{3pt}
\begin{tcolorbox}[colback=gray!20, colframe=black!80,
  title=\textbf{L3 (semantic inversion)}, fonttitle=\small\bfseries, boxrule=0.5pt]
\begin{lstlisting}[language=Python,
  basicstyle=\ttfamily\footnotesize,
  columns=fixed,keepspaces=true,tabsize=4]
def fib(n):
    if n <= 1:
        return fib(n-1) + fib(n-2)
    return n
\end{lstlisting}
\tcblower
\small\textit{Inverted: if-block runs when condition is FALSE}
\end{tcolorbox}
\caption{Semantic inversion in L3. Surface syntax is identical to Python, but if-block semantics are inverted. The correct L3 solution (right) looks counter-intuitive to Python-trained models, which generate the left version regardless of the stated rule.}
\label{fig:semantic_inversion}
\end{figure}

\subsection{L4: Semantic Inversion (Implicit, Example-Only)}
Same as L3, but no rule is stated. Instead, models receive 2--3 worked examples that demonstrate the inversion. Models must extract the semantic pattern through few-shot reasoning.

\subsection{Evaluation Metrics}
Redundant metrics are removed. We report two independent measures:
\begin{itemize}
  \item \textbf{KLR} (Keyword Leakage Rate): fraction of alias slots where original Python keyword was used. Note: ACR $= 1 - \text{KLR}$ and PIR $\approx \text{KLR}$ are derived quantities and not separately reported.
  \item \textbf{PSS} (Partial Structural Score, exploratory): weighted sub-component credit (alias: 0.4, parse: 0.25, skeleton: 0.25, format: 0.1). Weights are heuristic; PSS is reported as a secondary diagnostic alongside KLR and should not be used as the sole basis for compliance conclusions.
  \item \textbf{PPR} (Python Prior Rate, L3/L4): fraction of samples using Python-style condition (no inversion).
\end{itemize}

\subsection{Threats to Validity}
\textbf{Internal}: judge rules may be stricter than human acceptance; PSS weights are heuristic.
\textbf{Construct}: multi-task extension (T1 fib, T2 is\_sorted, T3 count\_vowels) partially addresses single-task limitation; broader coverage (nested control-flow, stateful logic) remains for future work. Preliminary open-weight evaluation (\texttt{Llama-3.3-70B}, \texttt{Qwen3-32B}) is reported at $n=20$, Variant~A only (§\ref{sec:open_weight}).
\textbf{External}: primary results cover nine model families (seven OpenAI + two open-weight via Groq); open-weight evaluation is pilot-scale ($n=20$, single variant) and full multi-variant replication remains future work. Human baseline (Exp-5) needed to validate the educational claim.
