\section{Experiments}
\subsection{Experimental Overview}
We ran experiments at each language level for available model families. All runs used prompts from \texttt{v1--v120} (L1) or dedicated level-specific banks (L2--L4). API authentication was managed via project environment variables.

\subsection{Models Evaluated}
\begin{itemize}
  \item \textbf{Core OpenAI} (seven models): \texttt{gpt-4o}, \texttt{gpt-4o-mini}, \texttt{gpt-4.1}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}, \texttt{gpt-5.4-mini}, \texttt{o4-mini}
  \item \textbf{Frontier extension}: \texttt{gpt-5.4}, \texttt{o3} (hard-task extension §\ref{sec:l4_hard} and E24 Fibonacci ablation §\ref{sec:l4_ablation})
  \item \textbf{Open-weight via Groq}: \llamaseventy{} ($n_\mathrm{eff}=25$, per-variant underpowered), \qwenthirtytwo{} ($n=50$, full replication)
\end{itemize}

\subsection{L1 Experiment: Token Alias Substitution}
120 prompts per model, versioned v1--v120 with varied alias mappings. Context-pack ablation: baseline vs. explicit alias contract system prompt.

\subsection{L2 Experiment: Syntax Inversion}
5 prompts $\times$ 4 runs each ($n=20$) using \texttt{:define ... ->} syntax (see Appendix). Model given complete grammar specification. Evaluated for \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano} (plus $n=5$ for \texttt{gpt-4o-mini} from earlier run).

\subsection{L3 Experiment: Semantic Inversion (Explicit)}
10 prompts $\times$ 2 runs each ($n=20$). System message explicitly states: ``if-blocks execute when condition is FALSE.'' Evaluated for \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano} (plus $n=5$--10 for \texttt{gpt-4o-mini} and \texttt{gpt-4o} from earlier runs).

\subsection{L4 Experiment: Semantic Inversion (Implicit)}
Initial screening: 10 prompts per model, no rule stated; only 2 worked examples with inverted semantics shown. Model must infer pattern.
Ablation (n=50): 5 prompt variants $\times$ 10 runs each, varying the number of worked examples (2 vs.\ 3), presence of inline comments, order of examples, and domain of examples. Evaluated across all seven models at $n=50$ each (350 total L4 ablation runs): \texttt{gpt-4o}, \texttt{gpt-4o-mini}, \texttt{gpt-4.1}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}, \texttt{gpt-5.4-mini}, and \texttt{o4-mini}.

\subsection{L4 Multi-task Generalization Experiment}
To assess whether pattern blindness generalizes across algorithmic domains, we extended the L4 evaluation to two additional tasks beyond Fibonacci:
\textbf{T2 (is\_sorted)}: write a function returning True if a list is ascending, and
\textbf{T3 (count\_vowels)}: count vowel characters in a string.
Both use a shared prompt header with two worked examples demonstrating inverted semantics plus inline code comments that \emph{partially} annotate the inversion rule (e.g., {\small\texttt{\# runs when FALSE}}). This is a \emph{partially-annotated L4} condition---more informative than the strict example-only Variant~A ablation (\S\ref{sec:l4_ablation}) and therefore \textbf{not directly comparable} to Table~\ref{tab:l4_ablation}.
We ran $n=10$ per task for \texttt{gpt-4o}, \texttt{gpt-4o-mini}, \texttt{gpt-4.1}, \texttt{gpt-4.1-nano}, and \texttt{gpt-5.4-mini}; $n=20$ for \texttt{gpt-4.1-mini}; and $n=15$ for \texttt{o4-mini} (255 initial runs). We then scaled \texttt{gpt-4o} T3 and \texttt{gpt-4.1} T2 to $n=50$ to resolve preliminary findings, yielding 345 total multi-task runs. Temperature=0 for all models except \texttt{o4-mini} (temperature=1, API requirement).
Results are in Table~\ref{tab:l4_multitask}.

\subsection{CoT Ablation Experiment}
To test whether chain-of-thought reasoning overcomes pattern blindness, we added a CoT instruction (``think step by step about the pattern first, then write code'') to the L4 Variant A prompt. We compared CoT vs.\ no-CoT for four models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}, and \texttt{o4-mini} ($n=20$ per condition, 160 total runs). Judge criteria are identical to the L4 ablation; we additionally track whether the model's reasoning text mentions the inversion pattern (operationally defined as containing at least one of the substrings: \texttt{"inverted"}, \texttt{"opposite"}, \texttt{"false when"}, or \texttt{"condition is false"}, case-insensitive regex match on the full reasoning output).

\subsection{L4 Hard-Task Extension: Complex Algorithms and Latest Models}
To test whether task complexity amplifies or attenuates pattern blindness, we designed three ``hard'' L4 tasks requiring multi-step algorithms with deep Python priors:
\textbf{H1 (merge\_sort)}: implement merge sort that returns \emph{descending} order (prior: ascending);
\textbf{H2 (binary\_search)}: return $-1$ if element is \emph{found}, insertion index if \emph{not found} (prior: index if found, $-1$ if not);
\textbf{H3 (bfs)}: return BFS traversal in \emph{reverse} order (prior: standard BFS order).
Each prompt provides 2 worked I/O examples demonstrating the inverted behavior, with no explicit rule.
We evaluated four models: \texttt{gpt-4o} (baseline), \texttt{gpt-5.4} (latest dense), \texttt{o3} (reasoning), and \texttt{o4-mini} (reasoning); $n=10$ per model per task (120 total runs).
Results are in Table~\ref{tab:l4_hard}.

\subsection{E24: Frontier Model L4 Fibonacci Ablation}
\label{sec:e24}
To directly address whether frontier models (\texttt{gpt-5.4}, \texttt{o3}) can overcome pattern blindness on prior-entrenched tasks---a gap in the hard-task extension---we ran the strict L4 Fibonacci ablation (Variant A, example-only, no explicit rule) on both models at $n=10$ each. This is the same Variant A condition used in Table~\ref{tab:l4_ablation}. Results are reported in \S\ref{sec:l4_ablation}.

\subsection{E28: Annotation Density Gradient (o4-mini, Pilot)}
\label{sec:e28}
\emph{Pilot for E34.} To map the annotation sensitivity boundary for reasoning models---motivated by the dissociation between CoT ($1/20$ pass) and partial annotation ($8/15$ pass) for \texttt{o4-mini}---we ran a density gradient experiment varying the fraction of inline annotation comments across the 8 worked examples in a T1 Fibonacci L4 prompt. Four density levels were tested: 25\%, 50\%, 75\%, and 100\% of examples annotated with inverted-semantics comments. $n=10$ per density level (40 total runs, \texttt{o4-mini} only). This pilot directly operationalizes the annotation-sensitivity hypothesis; the full multi-model replication is E34 (\S\ref{sec:e34}).

\subsection{E25--E27: Diagnostic L3 for \texttt{gpt-4.1} and \texttt{Qwen3-32B}}
\label{sec:e25e27}
E25: To determine whether \texttt{gpt-4.1}'s L4 T2 failure (1/50) reflects incapacity for semantic inversion or purely pattern-extraction failure, we ran L3 T2 evaluations (\texttt{is\_sorted}, explicit rule) for \texttt{gpt-4.1} at $n=10$. E26/E27: To determine whether \texttt{Qwen3-32B}'s L4 failure is L4-specific or reflects a deeper L3 instruction-following deficit, we ran L3 Fibonacci evaluations at $n=5$ (E26 pilot) and $n=20$ full replication (E27). Results are in \S\ref{sec:nonmonotonic} and \S\ref{sec:open_weight_discussion}.

\subsection{E31: Prior Entrenchment Score via Logprobs}
\label{sec:e31}
To directly quantify prior strength at the key generation decision point, we queried the logprobs API (\texttt{top\_logprobs=5}, \texttt{max\_tokens=1}, \texttt{temperature=0}) at the token immediately following \texttt{return} for four tasks (fib, merge\_sort, is\_sorted, count\_vowels) across two models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}). PES is defined as the probability assigned to the canonical Python token at this position. This directly measures the prior that L4 must overcome. Results are in \S\ref{sec:e31_pes}.

\subsection{E32: Cross-Task Transfer}
\label{sec:e32}
To test whether L4 resistance survives cross-task contamination, we evaluated L4 Fibonacci pass rate after showing inverted-semantics code examples from two unrelated tasks (merge sort, binary search) in the same prompt. \texttt{gpt-4.1-mini} and \texttt{o4-mini}: $n=10$; \texttt{gpt-4o}: scaled to $n=30$. Control condition (E32-ctrl): normal (non-inverted) sort examples preceding fib, \texttt{gpt-4o}, $n=10$. Temperature=0.7 for all. Results are in \S\ref{sec:e32_transfer}.

\subsection{E33 and E35: I/O Transparency Experiments}
\label{sec:e33e35}
E33 tested IO-A (transparent Fibonacci-adjacent mapping) vs.\ IO-B (opaque counter function with disclosed algorithm), revealing a confound: IO-B disclosed algorithmic steps, making both tasks procedurally transparent. Models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4o-mini}) achieved 10/10 on both. E33v2 replicated with \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{o4-mini} to confirm confound robustness. E35 (redesigned) replaced IO-B with pure I/O pairs (8 examples, no algorithm description) to cleanly isolate opacity as an independent variable. $n=10$ per cell, temperature=0.7. Results are in \S\ref{sec:e33_io}.

\subsection{E34: Annotation Density Curve (Multi-Model, Full Replication)}
\label{sec:e34}
Full replication of the E28 pilot (\S\ref{sec:e28}), extending to three models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{o4-mini}) across five density levels (0\%, 25\%, 50\%, 75\%, 100\%) on L4 Fibonacci with inverted comparisons. The 0\% condition includes an inverted \emph{code} example (not I/O-only), making it qualitatively distinct from verbal annotation levels (25--100\%). $n=10$ per cell (150 total runs). Results are in \S\ref{sec:e34_density}.

\subsection{Judge Validation}
To validate judge accuracy, we manually inspected a 50-run random sample across three models and three variants. The automated judge agreed with manual assessment in 48/50 cases (96\% agreement); the 2 disagreements involved partial inversion outputs conservatively classified as fail. Full validation data is in the repository.

\subsection{Reproducibility}
All experimental artifacts, commands, and results are committed to the repository under \texttt{docs/research/results/}. See reproducibility protocol section for commands.
