\section{Results}
\label{sec:results}

\begin{table}[!t]
\centering
\caption{L1 evaluation: Keyword Leakage Rate (KLR) per model. Baseline rows use $n=120$; ctx-pack rows use $n=20$. Models sorted by KLR (ascending). Context-pack reduces KLR for paired models: \texttt{gpt-4o-mini} ($-38\%$), \texttt{gpt-4.1-mini} ($-51\%$). Open-weight models (marked $^\star$) via Groq API, ctx-pack only. PSS (heuristic diagnostic) reported in Appendix~\ref{sec:appendix_pss}.}
\label{tab:l1_results}
\footnotesize
\begin{tabular}{@{}llc@{}}
\toprule
\textbf{Model} & \textbf{Delivery} & \textbf{KLR}$\downarrow$ \\
\midrule
\texttt{gpt-4.1-mini} & ctx-pack & \cellcolor{green!15}\textbf{0.21} \\
\texttt{Llama-3.3-70B}$^\star$ & ctx-pack & \cellcolor{green!15}0.23 \\
\texttt{gpt-4o-mini}  & ctx-pack & \cellcolor{green!15}0.26 \\
\texttt{o4-mini}      & ctx-pack & \cellcolor{green!15}0.27 \\
\texttt{gpt-4.1}      & ctx-pack & 0.30 \\
\texttt{gpt-4o}       & baseline & 0.37 \\
\texttt{gpt-4.1-nano} & ctx-pack & 0.38 \\
\texttt{gpt-4o-mini}  & baseline & 0.42 \\
\texttt{gpt-4.1-mini} & baseline & 0.43 \\
\texttt{gpt-5.4-mini} & baseline & \cellcolor{red!10}0.67 \\
\texttt{gpt-4.1-nano} & baseline & \cellcolor{red!10}0.69 \\
\texttt{Qwen3-32B}$^\star$ & ctx-pack & \cellcolor{red!10}1.00 \\
\bottomrule
\multicolumn{3}{l}{\scriptsize $^\star$Open-weight via Groq API; $n=20$, ctx-pack only. See \S\ref{sec:open_weight}.} \\
\multicolumn{3}{l}{\scriptsize \colorbox{green!15}{Green}: KLR${\leq}0.27$;\; \colorbox{red!10}{Red}: KLR${\geq}0.67$.} \\
\end{tabular}
\end{table}

\begin{figure}[!t]
\centering
\begin{tikzpicture}[font=\small, x=4.2cm]
  \def\barh{0.28}
  \def\barsep{0.50}
  % grid lines
  \foreach \x in {0, 0.25, 0.5, 0.75, 1.0} {
    \draw[dotted,gray!45] (\x,{-0.1}) -- (\x, {6.5*\barsep});
    \node[below, font=\tiny] at (\x, {-0.1}) {\x};
  }
  % 50% threshold
  \draw[dashed, red!60!black, line width=0.8pt] (0.5,{-0.15}) -- (0.5,{6.5*\barsep})
    node[above, font=\tiny\bfseries, text=red!60!black] {50\%};
  % row 0: gpt-4o
  \fill[blue!60] (0,{0*\barsep}) rectangle (0.37,{0*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{0*\barsep+\barh/2}) {gpt-4o};
  \node[right,font=\tiny,xshift=2pt] at (0.37,{0*\barsep+\barh/2}) {0.37};
  % row 1: gpt-4o-mini
  \fill[blue!28] (0,{1*\barsep}) rectangle (0.26,{1*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{1*\barsep+\barh/2}) {gpt-4o-mini};
  \node[right,font=\tiny,xshift=2pt] at (0.26,{1*\barsep+\barh/2}) {0.26};
  % row 2: gpt-4.1
  \fill[orange!65] (0,{2*\barsep}) rectangle (0.30,{2*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{2*\barsep+\barh/2}) {gpt-4.1};
  \node[right,font=\tiny,xshift=2pt] at (0.30,{2*\barsep+\barh/2}) {0.30};
  % row 3: gpt-4.1-mini
  \fill[orange!42] (0,{3*\barsep}) rectangle (0.21,{3*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{3*\barsep+\barh/2}) {gpt-4.1-mini};
  \node[right,font=\tiny,xshift=2pt] at (0.21,{3*\barsep+\barh/2}) {0.21};
  % row 4: gpt-4.1-nano
  \fill[orange!22] (0,{4*\barsep}) rectangle (0.38,{4*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{4*\barsep+\barh/2}) {gpt-4.1-nano};
  \node[right,font=\tiny,xshift=2pt] at (0.38,{4*\barsep+\barh/2}) {0.38};
  % row 5: gpt-5.4-mini
  \fill[green!50] (0,{5*\barsep}) rectangle (0.67,{5*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{5*\barsep+\barh/2}) {gpt-5.4-mini};
  \node[right,font=\tiny,xshift=2pt] at (0.67,{5*\barsep+\barh/2}) {0.67};
  % row 6: o4-mini
  \fill[violet!55] (0,{6*\barsep}) rectangle (0.27,{6*\barsep+\barh});
  \node[left,font=\scriptsize\ttfamily,xshift=-3pt] at (0,{6*\barsep+\barh/2}) {o4-mini};
  \node[right,font=\tiny,xshift=2pt] at (0.27,{6*\barsep+\barh/2}) {0.27};
  % x-axis label
  \node[below=8pt, font=\small] at (0.5,{-0.1}) {KLR (lower = better)};
\end{tikzpicture}
\caption{L1 KLR per model (OpenAI models only). Ctx-pack condition shown for all models except \texttt{gpt-4o} and \texttt{gpt-5.4-mini} (baseline only; ctx-pack not evaluated). $n{\geq}20$ per bar. Open-weight models (\texttt{Llama-3.3-70B}, \texttt{Qwen3-32B}) appear in Table~\ref{tab:l1_results} only; bar chart shows OpenAI models for visual clarity.
  \textcolor{blue!65}{\rule{5pt}{5pt}}~\texttt{gpt-4o} fam.;
  \textcolor{orange!60}{\rule{5pt}{5pt}}~\texttt{gpt-4.1} fam.;
  \textcolor{green!55!black}{\rule{5pt}{5pt}}~\texttt{gpt-5.4-mini} (baseline);
  \textcolor{violet!60}{\rule{5pt}{5pt}}~\texttt{o4-mini}.}
\label{fig:l1_klr_bar}
\end{figure}

\subsection{L1: Partial resistance; model-family compliance patterns}
All OpenAI models show non-zero KLR (0.21--0.69), confirming prior dominance. Three findings stand out.

(1) \texttt{gpt-5.4-mini} (KLR=0.67) underperforms \texttt{gpt-4o} (KLR=0.37) despite being a newer model; because these models differ in training corpus and RLHF regime---not just scale---we characterize this as \emph{model-family-specific compliance patterns} rather than a scale effect. The observed difference may reflect stronger code-generation fine-tuning in \texttt{gpt-5.4-mini} (a stated hypothesis, not a measured mechanism).

(2) For models with paired conditions, context-pack delivery reduces KLR: \texttt{gpt-4o-mini} ($0.42 \to 0.26$, $-38\%$) and \texttt{gpt-4.1-mini} ($0.43 \to 0.21$, $-51\%$).

(3) The reasoning model \texttt{o4-mini} (KLR=0.27) performs comparably to \texttt{gpt-4o-mini} ctx-pack (0.26), suggesting chain-of-thought reasoning provides modest benefit for alias compliance.

\noindent\textit{Note:} \texttt{gpt-4.1}, \texttt{gpt-4.1-nano}, and \texttt{o4-mini} were only evaluated under ctx-pack ($n=20$, v1--v20); baseline comparison for these models is not available and should be interpreted with caution given the smaller and non-overlapping prompt subset.

\textbf{Open-weight preliminary:} \texttt{Llama-3.3-70B} (KLR=0.23) is comparable to the best OpenAI ctx-pack results; \texttt{Qwen3-32B} (KLR=1.00) is the worst result across all tested models. This cross-family variance (0.23 vs.\ 1.00) exceeds the spread within the \texttt{gpt-4.x} family, indicating that architecture and instruction-following fine-tuning---not parameter count---govern alias compliance. See \S\ref{sec:open_weight} for full details and limitations.

\begin{table}[!t]
\centering
\caption{L2--L4 outcomes by model and condition. L2 is trivially adopted by all models (three at $n=20$; \texttt{gpt-4o-mini} pilot $n=5$); L3 reveals a model-capability threshold (\texttt{gpt-4o-mini} fails, others succeed); L4 (example-only) defeats all models despite L3 success. Rows with $n=5$ are pilot results; Wilson 95\% CIs shown.}
\label{tab:l234_results}
\footnotesize
\begin{tabular}{@{}lllcc@{}}
\toprule
\textbf{Model} & \textbf{Level} & \textbf{Delivery} & \textbf{Pass} & \textbf{Failure mode} \\
\midrule
\texttt{gpt-4o-mini} & L2 & Expl.\ grammar & \cellcolor{green!15}5/5$^\dagger$   & None (adopted) \\
\texttt{gpt-4o}      & L2 & Expl.\ grammar & \cellcolor{green!15}20/20 & None (adopted) \\
\texttt{gpt-4.1-mini}& L2 & Expl.\ grammar & \cellcolor{green!15}20/20 & None (adopted) \\
\texttt{gpt-4.1-nano}& L2 & Expl.\ grammar & \cellcolor{green!15}20/20 & None (adopted) \\
\midrule
\texttt{gpt-4o-mini} & L3 & Expl.\ rule    & \cellcolor{red!10}0/5$^\dagger$   & Prior dominance \\
\texttt{gpt-4o}      & L3 & Expl.\ rule    & \cellcolor{green!15}20/20 & --- (success) \\
\texttt{gpt-4.1-mini}& L3 & Expl.\ rule    & \cellcolor{green!15}20/20 & --- (success) \\
\texttt{gpt-4.1-nano}& L3 & Expl.\ rule    & \cellcolor{green!15}20/20 & --- (success) \\
\midrule
\texttt{gpt-4o}      & L4 & Examples only & \cellcolor{red!10}0/5$^\dagger$   & Pattern blindness \\
\texttt{gpt-4.1-nano}& L4 & Examples only & \cellcolor{red!10}0/50  & Pattern blindness \\
\bottomrule
\multicolumn{5}{l}{\scriptsize $^\dagger$Pilot ($n=5$); Wilson 95\% CI: [57\%,100\%] for 5/5, [0\%,43\%] for 0/5.} \\
\end{tabular}
\end{table}

\subsection{L2: Syntax inversion not resistant}
Three models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}) adopted L2 syntax fully at $n=20$ each (SIR=0.0; Wilson 95\% CI: [84\%,100\%] for adoption rate); \texttt{gpt-4o-mini} adopted L2 in a pilot evaluation ($n=5$, 5/5; Wilson CI: [57\%,100\%]). Across all four evaluated models, SIR=0.0, confirming that structural novelty alone does not create LLM resistance.

\subsection{L3: Capability threshold at semantic rule-following}
With an explicit semantic inversion rule, \texttt{gpt-4o-mini} fails completely (0/5, PPR=1.0): it generates standard Python despite instruction. By contrast, \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, and \texttt{gpt-4.1-nano} all succeed (20/20, PPR=0.0). The gap between \texttt{gpt-4o-mini} (0/5) and \texttt{gpt-4o} (20/20) is significant (Fisher's exact $p < 0.0001$), revealing a model capacity threshold for following semantic meta-rules.

\textbf{Key contrast: L3 success $\nRightarrow$ L4 success.} \texttt{gpt-4.1-nano} achieves 20/20 at L3 but 0/50 (PPR=1.0) at L4---the same model that correctly follows an explicit inversion rule is completely blind to the same rule embedded only in examples. This within-model contrast directly isolates pattern blindness as a failure mode independent of model capacity to follow rules when stated explicitly.

\subsection{L4: Pattern blindness across all tested models}
\label{sec:l4_ablation}
Without explicit rule, \texttt{gpt-4o}---which succeeds at L3---fails at L4 (PPR=1.0), producing canonical Python and ignoring example-embedded inversion cues. Figure~\ref{fig:l4_failure} shows the contrast.

\textbf{L4 ablation (n=50 per model).}
We scaled the L4 evaluation across seven models, all at $n=50$ (\texttt{gpt-4o}, \texttt{gpt-4o-mini}, \texttt{gpt-4.1}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}, \texttt{gpt-5.4-mini}, \texttt{o4-mini}).
Results are shown in Table~\ref{tab:l4_ablation}.
All seven models achieve 0\% pass rate, confirming that pattern blindness is robust across model families and including the reasoning model \texttt{o4-mini}. (\texttt{gpt-5.4-mini}: 0/42 valid runs, $n_\mathrm{eff}=42$; see footnote.)
PPR values range from 0.38 (\texttt{gpt-4.1-mini}) to 1.0 (\texttt{gpt-4.1}, \texttt{gpt-4.1-nano}, \texttt{gpt-5.4-mini}); \texttt{o4-mini} shows PPR=0.92. (\texttt{gpt-4.1-mini}'s 0.38 overall PPR reflects variant heterogeneity, not a uniformly lower prior pull; variant-level PPR spans 0.00--1.00, with certain variants achieving full inversion while others revert completely---see Table~\ref{tab:l4_ablation} for per-variant values.)
Notably, \texttt{gpt-4.1} and \texttt{gpt-4.1-nano} show PPR=1.0 across \emph{all} 50 runs and all variants, indicating complete prior dominance with no evidence of inversion detection.
The \texttt{o4-mini} result revises an earlier preliminary finding (1/20 at $n=20$): at the full scale of $n=50$, \texttt{o4-mini} also achieves 0/50, indicating that chain-of-thought reasoning does not provide a reliable pathway to overcoming pattern blindness at this scale.

\textbf{E24: Frontier models on L4 Fibonacci (strict example-only, $n=10$).}
To close the gap identified in \S\ref{sec:l4_hard}---where \texttt{gpt-5.4} and \texttt{o3} were only tested on hard tasks---we ran both frontier models on the identical Variant~A condition used in Table~\ref{tab:l4_ablation} (strict example-only, Fibonacci, $n=10$). Result: \texttt{gpt-5.4} 0/10 (PPR=1.00) and \texttt{o3} 0/10 (PPR=1.00). Both frontier models show complete prior dominance on L4 Fibonacci, confirming that pattern blindness persists even in the latest dense and reasoning models when the task has a deeply entrenched Python prior. This result directly confirms the core claim: \emph{L4 pattern blindness on prior-entrenched tasks is not defeated by frontier models as of 2026-03-25}.

\begin{table}[!t]
\centering
\caption{L4 ablation (PPR = Python Prior Rate). Seven core models: $n=50$ each (5 variants $\times$ 10 runs); \texttt{gpt-5.4-mini} $n_\mathrm{eff}=42$ (8 HTTP errors in Var.~E, excluded). Pass rate = 0 for all models. \texttt{o4-mini} revises $n=20$ preliminary (1/20). $^\S$E24 frontier extension: \texttt{gpt-5.4} and \texttt{o3} also tested at Var.~A, $n=10$: both 0/10 pass, PPR=1.00.}
\label{tab:l4_ablation}
\footnotesize
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}llccccccc@{}}
\toprule
\textbf{Var.} & \textbf{Description} & \makecell[c]{\small\texttt{gpt-}\\\small\texttt{4o}} & \makecell[c]{\small\texttt{gpt-4o-}\\\small\texttt{mini}} & \makecell[c]{\small\texttt{gpt-}\\\small\texttt{4.1}} & \makecell[c]{\small\texttt{gpt-4.1-}\\\small\texttt{mini}} & \makecell[c]{\small\texttt{gpt-4.1-}\\\small\texttt{nano}} & \makecell[c]{\small\texttt{gpt-5.4-}\\\small\texttt{mini}} & \makecell[c]{\small\texttt{o4-}\\\small\texttt{mini}} \\
\midrule
A & 2ex-baseline     & 0.40 & \textbf{0.10} & 1.00 & \textbf{0.10} & 1.00 & 1.00 & 0.80 \\
B & 3ex-extra        & 1.00 & 1.00 & 1.00 & \textbf{0.00} & 1.00 & 1.00 & 1.00 \\
C & 2ex-no-comment   & 0.90 & 1.00 & 1.00 & 0.80 & 1.00 & 1.00 & 0.90 \\
D & 2ex-swapped      & 0.70 & 1.00 & 1.00 & \textbf{0.00} & 1.00 & 1.00 & 1.00 \\
E & 3ex-diff-domain  & 0.50 & 1.00 & 1.00 & 1.00 & 1.00 & N/A$^\dagger$ & 0.90 \\
\midrule
\multicolumn{2}{l}{\textbf{Overall PPR}} & 0.70 & 0.82 & 1.00 & \textbf{0.38} & 1.00 & 1.00$^\ddagger$ & 0.92 \\
\multicolumn{2}{l}{\textbf{Pass rate}} & \cellcolor{red!10}0/50 & \cellcolor{red!10}0/50 & \cellcolor{red!10}0/50 & \cellcolor{red!10}0/50 & \cellcolor{red!10}0/50 & \cellcolor{red!10}0/42$^\dagger$ & \cellcolor{red!10}0/50 \\
\bottomrule
\multicolumn{9}{l}{\scriptsize $^\dagger$\texttt{gpt-5.4-mini} var-E: 8 HTTP errors; $n_\mathrm{eff}=2$; excluded from per-variant PPR.} \\
\multicolumn{9}{l}{\scriptsize $^\ddagger$Overall PPR for \texttt{gpt-5.4-mini} computed over 4 valid variants (A--D); var-E excluded.} \\
\multicolumn{9}{l}{\scriptsize $^\S$E24: \texttt{gpt-5.4} 0/10 (PPR=1.00), \texttt{o3} 0/10 (PPR=1.00) on Var.~A only; see text.} \\
\end{tabular}}
\end{table}

\noindent\textbf{Variant E domain-shift anomaly (\texttt{gpt-4.1-mini}).}
The most striking per-variant result is \texttt{gpt-4.1-mini}'s complete inversion across variant E: PPR=0.00 on variants~B (3ex-extra) and~D (2ex-swapped)---meaning it never generates the Python prior---yet PPR=1.00 on variant~E (3ex-diff-domain, where examples use a \emph{different} semantic domain). The structural difference is that variants B and D use the same Fibonacci-adjacent domain for all examples, while variant E uses cross-domain examples (e.g., sorting or arithmetic comparisons). We hypothesize that \emph{domain shift in examples weakens the model's pattern extraction signal}: when examples share the same function context as the target task, the model's in-context pattern extractor can align examples to the task template; when examples are drawn from a different domain, the learned pattern loses its grounding in the target task structure, causing the model to revert to the Python prior. This remains a hypothesis; controlled cross-domain vs.\ in-domain replication at larger $n$ would be needed to confirm it. Notably, variant~A (2ex-baseline, in-domain) achieves PPR=0.10 while variant~B (3ex, in-domain) achieves PPR=0.00: example count and domain \emph{jointly} modulate extraction quality in this model---additional in-domain examples may partially suppress the Python prior without enabling full inversion, while domain shift reinvigorates it.

\begin{figure}[!t]
\centering
\begin{tcolorbox}[colback=gray!8, colframe=black!60,
  title=\textbf{{\checkmark} Panel A: L3 --- explicit rule (\texttt{gpt-4o}) [CORRECT]},
  fonttitle=\small\bfseries, boxrule=0.5pt]
\textit{System}: ``if-blocks run when condition is FALSE.''\\[3pt]
\textit{\texttt{gpt-4o} output} \textbf{(CORRECT):}\\[2pt]
\begin{lstlisting}[language=Python,
  basicstyle=\ttfamily\footnotesize,
  columns=fixed,keepspaces=true,tabsize=4,
  commentstyle=\color{green!60!black}]
if n > 1:        # inverted
    return n
return fib(n-1) + fib(n-2)
\end{lstlisting}
\end{tcolorbox}
\vspace{3pt}
\begin{tcolorbox}[colback=gray!20, colframe=black!80,
  title=\textbf{{\texttimes} Panel B: L4 --- examples only (\texttt{gpt-4o}) [INCORRECT]},
  fonttitle=\small\bfseries, boxrule=0.5pt]
\textit{User}: [2 worked examples with inverted semantics, no explicit rule]\\[3pt]
\textit{\texttt{gpt-4o} output} \textbf{(INCORRECT):}\\[2pt]
\begin{lstlisting}[language=Python,
  basicstyle=\ttfamily\footnotesize,
  columns=fixed,keepspaces=true,tabsize=4,
  commentstyle=\color{red!70!black}]
if n <= 1:       # standard Python
    return n
return fib(n-1) + fib(n-2)
\end{lstlisting}
\end{tcolorbox}
\caption{Prior dominance vs.\ pattern blindness isolated on the same model (\texttt{gpt-4o}). \textbf{A}: \texttt{gpt-4o} correctly inverts logic with explicit rule (L3). \textbf{B}: The same model reverts to Python prior when the rule is embedded only in examples (L4). This contrast isolates pattern blindness as a failure mode distinct from prior dominance.}
\label{fig:l4_failure}
\end{figure}

\subsection{L4 Multi-task Generalization}
\label{sec:l4_multitask}
\textbf{Protocol note:} This section uses \emph{partially-annotated} prompts (inline comments partially state the inversion rule), which differ from the strict \emph{example-only} prompts in \S\ref{sec:l4_ablation}. Results in Table~\ref{tab:l4_multitask} are \textbf{not directly comparable} to Table~\ref{tab:l4_ablation}; comparisons across these sections should account for this prompt protocol difference.

To assess whether pattern blindness extends beyond Fibonacci, we evaluated all seven models on three tasks: T1 (\texttt{fibonacci}), T2 (\texttt{is\_sorted}), and T3 (\texttt{count\_vowels}). The multi-task prompt uses inline code comments that partially annotate the inversion rule (\textit{e.g.}, \texttt{\# runs when a > b is FALSE, i.e., a <= b}); this is a \emph{partially-annotated L4} condition---less explicit than L3 (no stated rule), but more informative than the pure example-only variants used in the ablation. \texttt{gpt-4o}, \texttt{gpt-4o-mini}, \texttt{gpt-5.4-mini}, and \texttt{o4-mini} were evaluated at $n=10$ or $n=15$ per task; \texttt{gpt-4.1-mini} at $n=20$ per task; \texttt{gpt-4.1} and \texttt{gpt-4.1-nano} at $n=10$ per task (255 initial runs). We further scaled \texttt{gpt-4o} T3 and \texttt{gpt-4.1} T2 to $n=50$ each to resolve preliminary ambiguities, and extended \texttt{o4-mini} T2/T3 to $n=20$ each, yielding 345 total multi-task runs. Results are in Table~\ref{tab:l4_multitask}.

\begin{table}[!t]
\centering
\caption{L4 multi-task summary (\textbf{345 runs total}). $\dagger$~Partially-annotated prompts; \textbf{NOT directly comparable to Table~\ref{tab:l4_ablation}}~(strict example-only). Green = pass, red = fail. Dominant failure mode per row in rightmost column. Full PPR values in Appendix~\ref{sec:appendix_l4_multitask_ppr}.}
\label{tab:l4_multitask}
\footnotesize
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}l ccc l@{}}
\toprule
\textbf{Model} & \textbf{T1: fib} & \textbf{T2: sorted} & \textbf{T3: vowels} & \textbf{Dominant failure} \\
\midrule
\texttt{gpt-4o}       & \cellcolor{red!10}0/10   & \cellcolor{green!15}\textbf{10/10} & \cellcolor{red!10}0/50$^a$  & T1: prior dom.; T3: op.\ sub.\ (94\%) \\
\texttt{gpt-4o-mini}  & \cellcolor{red!10}0/10$^b$& \cellcolor{red!10}0/10            & \cellcolor{red!10}0/10      & template fragmentation (T1); prior dom. \\
\texttt{gpt-4.1}      & \cellcolor{red!10}0/10   & \cellcolor{red!10}1/50$^c$         & \cellcolor{red!10}0/10      & prior dominance \\
\texttt{gpt-4.1-mini} & \cellcolor{red!10}0/20   & \cellcolor{green!15}\textbf{20/20} & \cellcolor{red!10}0/20      & T1: prior dom.; T2: extractable \\
\texttt{gpt-4.1-nano} & \cellcolor{red!10}0/10   & \cellcolor{red!10}0/10             & \cellcolor{red!10}0/10      & prior dominance \\
\texttt{gpt-5.4-mini} & \cellcolor{red!10}0/10   & \cellcolor{red!10}0/10             & \cellcolor{red!10}0/10      & prior dominance (all tasks) \\
\texttt{o4-mini}      & \cellcolor{green!15}\textbf{8/15}$^d$ & \cellcolor{red!10}0/20 & \cellcolor{red!10}0/20 & T1: partial extraction (reasoning); T2/T3: prior dom. \\
\bottomrule
\multicolumn{5}{l}{\scriptsize $^a$OpSub 47/50 (94\%, CI: [83\%,99\%]).\ $^b$Template fragmentation.} \\
\multicolumn{5}{l}{\scriptsize $^c$$n=50$ scale-up; Wilson CI: pass [0\%,10.5\%]; Fisher's $p{<}0.0001$ vs.\ \texttt{gpt-4.1-mini}.} \\
\multicolumn{5}{l}{\scriptsize $^d$Token budget corrected (400$\to$5000 tokens); see \S\ref{sec:l4_ablation} for details.} \\
\end{tabular}}
\end{table}

Seven findings emerge.

\smallskip\noindent
(1) \textbf{T2 is extractable only by mid-to-large models}: \texttt{gpt-4o} (10/10) and \texttt{gpt-4.1-mini} (20/20) pass T2 fully, while \texttt{gpt-4.1} (1/50, PPR=0.42), \texttt{gpt-4.1-nano} (0/10), \texttt{gpt-5.4-mini} (0/10), and \texttt{gpt-4o-mini} (0/10) fail. This reveals a non-monotonic capacity threshold: T2 extractability does not scale linearly within or across model families.

\smallskip\noindent
(2) \textbf{T1 (Fibonacci) is resistant for all non-reasoning models}: six of seven models fail T1, with high or full PPR (PPR=1.0 for \texttt{gpt-4o}, \texttt{gpt-4.1}, \texttt{gpt-4.1-mini}, \texttt{gpt-5.4-mini}; PPR=0.2 for \texttt{gpt-4.1-nano}; template fragmentation for \texttt{gpt-4o-mini}). The reasoning model \texttt{o4-mini} provides a partial exception under the partially-annotated prompt (see Finding~5 for detail).

\smallskip\noindent
(3) \textbf{Operational substitution dominates T3}: the scale-up (\texttt{gpt-4o} T3, $n=50$) reveals operational substitution in 47/50 runs (94\%; Wilson 95\% CI: [83\%, 99\%])---the model applies arithmetic transformations (\texttt{count -= 1; return -count}) to achieve correct output without adopting the L4 semantic rule. The preliminary $n=10$ result (1 pass) was a statistical artifact; at $n=50$, \texttt{gpt-4o} achieves 0/50 pass (Wilson 95\% CI: [0\%, 7.1\%]). \texttt{gpt-5.4-mini} T3 also exhibits operational substitution (PPR=0.0, all 10 runs), confirming this as a structurally distinct, frequent failure mode.

\smallskip\noindent
(4) \textbf{Model capacity determines T2 extractability}: \texttt{gpt-4o-mini}, \texttt{gpt-4.1}, \texttt{gpt-4.1-nano}, and \texttt{gpt-5.4-mini} all fail T2, while \texttt{gpt-4o} and \texttt{gpt-4.1-mini} succeed---consistent with the capacity threshold observed at L3.

\smallskip\noindent
(5) \textbf{Reasoning partially mitigates pattern blindness under partial annotation}: \texttt{o4-mini} achieves 8/15 on T1 under the partially-annotated prompt (after token budget correction), but fails T2 and T3 (0/20 each at extended $n=20$; overall 8/55 across T1+T2+T3). Under the strict example-only ablation (§~\ref{sec:l4_ablation}), \texttt{o4-mini} achieves 0/50---chain-of-thought reasoning alone cannot overcome pure pattern blindness, but reasoning models can exploit partial annotation cues that dense models miss.

\textbf{Methodological note on \texttt{o4-mini} (token budget correction).} An initial \texttt{o4-mini} T1 run used \texttt{max\_completion\_tokens}=400, which truncated the model's chain-of-thought before it could complete the inversion task; this yielded 0/15 pass (consistent with pattern blindness). Upon correcting the budget to \texttt{max\_completion\_tokens}=5000, \texttt{o4-mini} achieved 8/15 pass---representing a finding revision. This correction reveals that reasoning models can exploit partial annotation cues inaccessible to dense models, but only when given sufficient token budget to complete the reasoning chain. The original 0/50 result in the L4 ablation (strict example-only, corrected budget) confirms that this effect is annotation-contingent: token budget alone does not overcome pure pattern blindness.

\smallskip\noindent
(6) \textbf{Within-family inconsistency is confirmed}: \texttt{gpt-4.1} T2 scale-up ($n=50$) confirms the capacity gap is systematic---1/50 pass (Wilson 95\% CI: [0\%, 10.5\%]) vs.\ \texttt{gpt-4.1-mini} 20/20 (Fisher's exact $p < 0.0001$). Fine-tuning or RLHF choices appear to play a larger role than raw scale in determining L4 pattern extraction capability.

\smallskip\noindent
(7) \textbf{\texttt{gpt-5.4-mini} shows uniform prior dominance across all tasks}: 0/30 with T1 PPR=1.0, T2 PPR=1.0, and T3 operational substitution pattern, consistent with its L4 ablation PPR=1.0 on Fibonacci variants.

\subsection{Three Failure Modes}
\begin{itemize}
  \item \textbf{Prior Dominance} (L1, L3-weak, L4-T2-mini, L4-T1-all): stated or implicit rule is overridden by pretrained token distribution.
  \item \textbf{Pattern Blindness} (L4-T1): model cannot extract the semantic rule from examples alone, even when the same model follows the rule when stated explicitly (L3). T1 (Fibonacci) shows universal pattern blindness across all models.
  \item \textbf{Operational Substitution} (L4-T3-\texttt{gpt-4o}, $n=50$): model compensates via arithmetic output transformation rather than adopting the semantic inversion rule, indicating goal awareness without rule internalization. Confirmed as dominant at 94\% of T3 \texttt{gpt-4o} runs (Wilson 95\% CI: [83\%, 99\%]).
\end{itemize}
We also observe \emph{template fragmentation} in \texttt{gpt-4o-mini} T1: the model partially disrupts its standard template but does not achieve semantic inversion, yielding an incorrect but structurally novel output. We treat this as a sub-variant of prior dominance rather than a distinct fourth mode, given the small sample ($n=10$); further study at larger $n$ may warrant reclassification.

\smallskip
Task difficulty interacts with model capacity to determine which failure mode dominates. T2 (is\_sorted) is extractable by mid-to-large models; T1 (Fibonacci) is universally resistant; T3 (count\_vowels) triggers operational substitution as the dominant escape strategy for capable models.

\textbf{Failure taxonomy across 419 failed L4 runs} (ablation + multitask T1; 427 total pool including 8 \texttt{o4-mini} T1 passes): Type-I (exact Python prior) accounts for ${\approx}83.1\%$ of failures, Type-II (partial attempt) for ${\approx}12.6\%$, and Type-III (operational substitution) for ${\approx}4.3\%$. Type-I dominates in \texttt{gpt-4.1} and \texttt{gpt-4.1-nano} (100\% of failures); Type-II concentrates in \texttt{gpt-4.1-mini} (44\%) and \texttt{gpt-4o-mini} (32\%); Type-III appears primarily in \texttt{gpt-4o} (25\% of its failures).

\noindent\textit{Population note:} The 427 pooled runs consist of 342 valid strict example-only ablation runs (all failures; \texttt{gpt-5.4-mini} $n_\mathrm{eff}=42$ due to 8 HTTP errors, counted at 342 not 350) and 85 multitask T1 runs (8 passes by \texttt{o4-mini}, 77 failures). Total failures: 419. These two sub-populations differ in prompt annotation level; the low overall Type-III rate (4.3\%) is consistent with strict ablation, where no partial annotation cues are available to trigger operational substitution. In the multi-task sub-population, Type-III is observed primarily in \texttt{gpt-4o} T3 at $n=50$ (94\% operational substitution rate), but that data is counted in T3 (not T1) and is therefore excluded from the 427 pooled taxonomy. A full split analysis per prompt protocol is deferred to future work.

\subsection{CoT Ablation: Reasoning About Inversion Is Not Sufficient}
\label{sec:cot_ablation}
We tested whether chain-of-thought (CoT) prompting overcomes pattern blindness by adding ``think step by step about the pattern first'' to the L4 Variant A prompt ($n=20$ per condition, four models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4.1-nano}, \texttt{o4-mini}).

\noindent\textbf{Result}: CoT does not reliably improve pass rate. Three of four models achieve 0/20 in both CoT and no-CoT conditions (Wilson 95\% CI: [0\%,16\%]); \texttt{o4-mini} achieves 1/20 under CoT (5\%; Wilson CI: [0.9\%,24\%]; Fisher's exact vs.\ 0/20 no-CoT: $p=1.0$, not significant). However, CoT consistently changes \emph{reasoning behavior} across all models:
\begin{itemize}
  \item \texttt{gpt-4o}: CoT reduces PPR from 1.00 to 0.15 and raises \emph{mentions\_inversion} to 100\%---the model explicitly \emph{describes} the inverted pattern but still generates standard Python code.
  \item \texttt{gpt-4.1-mini}: CoT reduces PPR from 0.40 to 0.00 with 100\% mentions\_inversion rate---same reasoning--generation dissociation.
  \item \texttt{gpt-4.1-nano}: CoT reduces PPR from 1.00 to 0.20 and raises \emph{mentions\_inversion} to 100\%---the nano model shows the strongest PPR reduction under CoT, yet all 20 outputs still fail to implement the inversion correctly.
  \item \texttt{o4-mini}: CoT reduces PPR from 0.95 to 0.65 and raises \emph{mentions\_inversion} to 35\%. This is the most notable reasoning-model CoT result: while dense models reach near-perfect \emph{mentions\_inversion} (100\%) yet still fail, o4-mini reaches only 35\% mentions at $n=20$, suggesting a dissociation between CoT-prompting and chain-of-thought use in the reasoning model. The no-CoT baseline (PPR=0.95 at Variant~A, $n=20$) is slightly higher than the overall L4 ablation PPR=0.92 for \texttt{o4-mini} (which averages all five variants); this is expected since different variants show different PPR values. The CoT comparison should be interpreted within Variant~A only.
\end{itemize}
This reveals a \emph{reasoning--generation dissociation} that generalizes across four model families: models can articulate the semantic inversion rule when prompted to reason, yet their code generation defaults to the Python prior regardless. Pattern blindness is thus not a failure to \emph{detect} the pattern but a failure to \emph{override} the prior during generation. The finding is robust: 160 total CoT runs (40 per model, 2 conditions) across dense and reasoning architectures yield zero reliable pass rate improvement.

\textbf{CoT vs.\ partial annotation for \texttt{o4-mini}.} A notable dissociation emerges when comparing the CoT ablation result to the multi-task result for \texttt{o4-mini}: CoT instruction on strict Variant~A yields 1/20 (5\%), while partial annotation (inline comment hints) without CoT instruction yields 8/15 (53\%, Table~\ref{tab:l4_multitask}). This suggests that for reasoning models, partial rule disclosure is more effective than meta-cognitive prompting alone: CoT helps dense models articulate the inversion pattern but does not supply the rule content; partial annotation provides actual rule content that reasoning models can exploit. The implication for resistance design is that partial annotations---not just implicit example delivery---are the key vulnerability for reasoning models.

\subsection{Open-Weight Models: Preliminary L1 and L4 Results}
\label{sec:open_weight}
To assess whether the failure modes generalize beyond the OpenAI model family, we conducted preliminary evaluations on two open-weight models via Groq API: \texttt{Llama-3.3-70B}-Versatile and \texttt{Qwen3-32B}.

\textbf{L1 (ctx-pack, $n=20$):} \texttt{Llama-3.3-70B} achieves KLR=0.23, nearly matching the OpenAI best (\texttt{gpt-4.1-mini} ctx-pack, 0.21). \texttt{Qwen3-32B} achieves KLR=1.00: all 20 outputs contained at least one native Python keyword with no valid code skeleton---the worst L1 result across all tested models. The inter-model gap (0.23 vs.\ 1.00) exceeds the intra-family spread of \texttt{gpt-4.x} models (0.21--0.69), indicating that instruction-following fine-tuning rather than parameter count governs alias compliance.

\textbf{L4 Variant A ($n=20$):} Both open-weight models fail. \texttt{Llama-3.3-70B}: 0/20 pass, PPR=0.05 (1/20 strict Python prior; remaining 19 outputs applied standard control flow without inversion). \texttt{Qwen3-32B}: 0/18 pass ($n_\mathrm{eff}=18$; 2 HTTP errors), PPR=1.00 (complete prior dominance). Neither model triggered inversion detection (noticed\_inversion=0). These results extend the pattern blindness finding beyond OpenAI models.

\emph{Limitations and extended runs:} Initial pilot ($n=20$, Variant~A only) confirmed pattern blindness for both open-weight models. Full 5-variant ablation ($n=50$ attempted) was subsequently run for both models. \texttt{Qwen3-32B} completed 50/50 valid runs (0 HTTP errors): 0/50 pass, PPR=1.00 across all variants---consistent with pilot results, confirming complete prior dominance. \texttt{Llama-3.3-70B} achieved 25/50 valid runs (25 HTTP errors due to Groq rate limiting; all five variants represented, $n_\mathrm{eff}=5$ per variant): 0/25 pass. Per-variant PPR values vary substantially: Var.~A=0.00, Var.~B=0.00, Var.~C=0.60, Var.~D=1.00, Var.~E=1.00 (overall PPR=0.52). At $n_\mathrm{eff}=5$ per variant, Wilson 95\% CIs are wide: PPR=0.00 has CI [0\%, 43\%]; PPR=0.60 has CI [17\%, 93\%]; PPR=1.00 has CI [57\%, 100\%]. These intervals overlap substantially---statistical claims about variant-level differences are not warranted at this sample size. This per-variant \emph{pattern} suggests a qualitatively different failure mode from both prior dominance (PPR$\approx$1.0) and Qwen's complete prior dominance: \texttt{Llama-3.3-70B} may apply standard control flow without semantic inversion (low PPR on Variants~A/B) but revert to strict Python prior on Variants~D/E (domain-shifted examples). This potentially represents a third open-weight failure mode---\emph{context-sensitive prior reversion}---but the per-variant pattern cannot be statistically confirmed at $n_\mathrm{eff}=5$; full replication is deferred to future work.

\textbf{E26/E27: L3 diagnostic for \texttt{Qwen3-32B} ($n=5$ pilot, $n=20$ replication).}
To disambiguate whether \texttt{Qwen3-32B}'s L4 failure (PPR=1.00) reflects (a) pattern extraction failure (\emph{L4-specific}, same as OpenAI models) or (b) general instruction-following failure for semantic meta-rules, we ran $n=5$ L3 evaluations (E26) and subsequently $n=20$ full replication (E27) on \texttt{Qwen3-32B} (explicit rule: ``if-blocks execute when the condition is FALSE,'' Fibonacci task).
E26 result ($n=5$): 0/5 pass, PPR=0.80 (1/5 showed a structurally distinct non-pass failure).
E27 result ($n=20$, 2026-03-25): 0/20 pass, PPR=0.40 (8/20 exact Python prior; 12/20 FAIL-other). No valid passes at $n_\mathrm{eff}=20$; Wilson 95\% CI: [0\%, 17\%]. The PPR=0.40 is \emph{lower} than the E26 pilot (0.80), reflecting that a substantial fraction of failures are structurally distinct from the pure Python prior---the model often produces internally inconsistent code that attempts partial rule application but fails, rather than simply returning the standard Fibonacci template. This is consistent with \emph{confused partial instruction-following}, not clean prior dominance.
Unlike \texttt{gpt-4.1} (E25: 10/10 on L3 T2), \texttt{Qwen3-32B} fails to apply the semantic inversion rule even when stated explicitly across $n=20$ trials---confirming \emph{general semantic meta-rule failure} as a systematic property, not a small-$n$ artifact. This represents a qualitatively different failure mode from all tested OpenAI models: \texttt{Qwen3-32B} cannot reliably follow semantic inversions via prompt instructions alone, regardless of delivery (explicit vs.\ implicit).

\subsection{E28: Annotation Density Gradient for \texttt{o4-mini}}
\label{sec:e28_results}

\begin{table}[!t]
\centering
\caption{E28: Annotation density gradient for \texttt{o4-mini} on T1 Fibonacci ($n=10$ per level, 40 total runs; $n_\mathrm{eff}=9$ at 100\% due to 1 HTTP error). PPR = prior-pull rate (fraction of \texttt{fail\_prior} among all non-pass runs). Pass rate and PPR are both reported.}
\label{tab:e28_density}
\footnotesize
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccccc@{}}
\toprule
\textbf{Density} & \textbf{Annotated/Total} & \textbf{Pass} & \textbf{Fail-Prior} & \textbf{Fail-Other} & \textbf{Pass Rate} & \textbf{PPR} \\
\midrule
25\%  & 2/8 & 3  & 1 & 6 & 0.30 & 0.10 \\
50\%  & 4/8 & 6  & 1 & 3 & \textbf{0.60} & 0.10 \\
75\%  & 6/8 & 2  & 5 & 3 & 0.20 & 0.50 \\
100\%$^\dagger$ & 8/8 & 1  & 2 & 6 & 0.11 & 0.22 \\
\midrule
\multicolumn{7}{l}{\scriptsize $^\dagger$$n_\mathrm{eff}=9$ (1 HTTP error); pass rate = 1/9.} \\
\bottomrule
\end{tabular}%
}
\end{table}

The E28 experiment reveals that annotation density has a \textbf{non-monotonic} effect on \texttt{o4-mini}'s ability to extract the L4 inversion pattern. Results are in Table~\ref{tab:e28_density}.

\textbf{Key finding---optimal density at 50\%}: Pass rate peaks at 50\% annotation density (6/10, PPR=0.10). Below this, at 25\% density, pass rate drops to 3/10---insufficient annotation signal for the reasoning model's chain-of-thought. Above 50\%, at 75\% annotation, pass rate collapses to 2/10 and PPR \emph{rises} sharply to 0.50---a paradoxical result where more annotation triggers stronger prior conflict. At 100\% density, pass rate is lowest (1/9 valid runs; 1 HTTP error), though PPR partially recovers to 0.22.

\textbf{Interpretation---annotation saturation induces prior conflict}: The non-monotonic curve is consistent with a two-competing-signal model. At low density (25\%), the reasoning model has insufficient annotation cues to construct the inversion rule: it either succeeds via independent pattern recognition (30\%) or fails via non-prior confusion (60\%). At 50\% density, annotation signal is sufficient for reasoning chain completion without overwhelming the Fibonacci prior. At 75\%+, the dense annotation makes the inversion semantics \emph{salient} to the model's representation, which simultaneously activates the conflicting Python prior---triggering more frequent prior-dominant failures (PPR=0.50). This suggests a retrieval-interference mechanism: annotation cues strong enough to activate the task schema also activate the competing prior schema.

\textbf{Design implication}: For prior-entrenched tasks, resistance via annotation density is robust in a narrow band. Over-annotated prompts are \emph{not} strictly harder; they may in fact partially self-defeat by activating the prior they target. Optimal confusion language design should calibrate annotation density to remain below the prior-activation threshold---approximately 50\% of examples for reasoning models on Fibonacci-class tasks.

\textbf{Connection to CoT--annotation dissociation}: The E28 results reinforce Finding~5's CoT--annotation gap. The prior multi-task result (8/15 pass at partial annotation) used a fixed partial-annotation regime that approximates the 50\% optimal; the CoT ablation (1/20 pass) added meta-cognitive instruction without annotation content. E28 maps the mechanism: annotation works by supplying inversion-rule content at a density that clears the reasoning threshold without triggering prior conflict. CoT alone cannot cross this boundary because it adds reasoning scaffolding but no new rule content.

\subsection{E31: Prior Entrenchment Score via Logprobs}
\label{sec:e31_pes}

To directly measure prior strength, we queried the logprobs API at the key decision point of each task's base case (the token immediately after \texttt{return}). For each model, we recorded the top-5 token probabilities and the rank/probability of the canonical Python token. Table~\ref{tab:e31_pes} summarizes the results.

\begin{table}[!t]
\centering
\caption{E31: Prior Entrenchment Score (PES) --- probability of the canonical Python token at the base-case decision point. Models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}. \texttt{gpt-4o-mini} results excluded (model emits markdown formatting tokens rather than code tokens at this position). PES = $P(\text{canonical token} \mid \text{prompt prefix})$.}
\label{tab:e31_pes}
\footnotesize
\begin{tabular}{@{}llcc@{}}
\toprule
\textbf{Task} & \textbf{Canonical token} & \textbf{gpt-4o} & \textbf{gpt-4.1-mini} \\
\midrule
fib (T1)         & \texttt{n}    & 0.9996 & 1.0000 \\
merge\_sort (H1) & \texttt{arr}  & 0.9999 & 1.0000 \\
is\_sorted (T2)  & \texttt{True} & 0.9736 & 0.9985 \\
count\_vowels (T3) & \texttt{s}  & 0.1606 & 1.0000 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key finding---PES alone does not predict L4 resistance.} Both \texttt{fib} and \texttt{merge\_sort} show near-identical PES ($\geq$0.9996), yet their L4 outcomes diverge dramatically: fib is universally resistant (0/350 pass in ablation), while merge\_sort is solvable by all tested models in the hard-task condition (Table~\ref{tab:l4_hard}). This dissociation rules out a simple ``prior strength $\to$ resistance'' mapping and supports the hypothesis that I/O specification transparency---not prior magnitude alone---is the critical moderating variable (\S\ref{sec:l4_hard}). The PES metric quantifies the \emph{strength} of the prior that must be overridden, but the model's ability to override it depends on whether the task's I/O examples make the target behavior transparent.

\texttt{is\_sorted} (T2) shows slightly lower PES (0.9736 for \texttt{gpt-4o}), which may partly explain its extractability by capable models---but PES=0.9736 is still very high, suggesting that the ${\sim}0.03$ PES gap is not the primary driver of the T1--T2 pass rate difference (0\% vs.\ 100\% for \texttt{gpt-4o}).

A secondary observation: \texttt{gpt-4o}'s low T3 PES (0.1606, \texttt{count\_vowels}) contrasts with its high operational substitution rate on T3 ($n=50$: 94\% OpSub). This suggests that a weaker prior (lower PES) does not simply yield L4 compliance; instead, the model explores alternative generation paths that still avoid semantic inversion, manifesting as arithmetic workarounds rather than canonical Python. The PES--operational-substitution relationship warrants future mechanistic study.

\subsection{E32: Cross-Task Transfer Breaks L4 Resistance}
\label{sec:e32_transfer}

A critical question for assessment design: can a model crack L4 Fibonacci if it first sees inverted semantics demonstrated on \emph{other} tasks? We tested few-shot cross-task transfer by providing two inverted code examples (merge\_sort with inverted comparisons, binary\_search with inverted comparisons) before requesting fib(n) in the same style. \texttt{gpt-4.1-mini} and \texttt{o4-mini} were evaluated at $n=10$; \texttt{gpt-4o} was scaled up to $n=30$ (two scale-up rounds) to tighten the confidence interval on its partial-transfer result. Temperature=0.7 for all. A \textbf{control condition} (E32-ctrl: ascending sort with \emph{normal}, non-inverted examples preceding fib, \texttt{gpt-4o}, $n=10$) was run to confirm that the transfer effect is specific to inversion semantics, not merely cross-task code exposure.

\begin{table}[!t]
\centering
\caption{E32: Cross-task transfer --- pass rate on L4 Fibonacci after exposure to inverted code from other tasks. Baseline: strict L4 example-only (Table~\ref{tab:l4_ablation}) yields 0/50 for all models. Wilson 95\% CIs shown. \texttt{gpt-4o} scaled to $n=30$ (E32a); others at $n=10$. E32-ctrl: control condition using normal (non-inverted) sort examples, \texttt{gpt-4o}, $n=10$.}
\label{tab:e32_transfer}
\footnotesize
\begin{tabular}{@{}lcccl@{}}
\toprule
\textbf{Model} & \textbf{Pass} & \textbf{Pass rate} & \textbf{Wilson 95\% CI} & \textbf{Note} \\
\midrule
\texttt{gpt-4.1-mini} & 10/10 & \cellcolor{green!15}\textbf{100\%} & [72\%,100\%] & Perfect transfer \\
\texttt{gpt-4o}       & 16/30 & \cellcolor{yellow!20}53\% & [36\%,70\%] & Partial transfer; vs.\ mini: $p{=}0.008^{**}$ \\
\texttt{o4-mini}      & 1/10  & \cellcolor{red!10}10\% & [2\%,44\%] & Minimal transfer \\
\midrule
\textit{E32-ctrl} (\texttt{gpt-4o}, normal examples) & 0/10 & \cellcolor{gray!10}0\% & [0\%,31\%] & No transfer (control) \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Cross-task contamination defeats L4.} The same models that achieve 0/50 under strict L4 example-only delivery (Table~\ref{tab:l4_ablation}) achieve 53--100\% pass when primed with inverted code from unrelated tasks. \texttt{gpt-4.1-mini} transfers perfectly (10/10; Wilson 95\% CI: [72\%,100\%]), consistent with its strong pattern extraction on T2 (20/20 in multi-task). \texttt{gpt-4o} achieves 16/30 at scale-up (Wilson 95\% CI: [36\%,70\%])---partial transfer confirmed; Fisher's exact vs.\ \texttt{gpt-4.1-mini}: $p=0.008$ ($p<0.01$), indicating a significant gap in transfer fidelity. \texttt{o4-mini}---which relies on reasoning chains rather than pattern mimicry---achieves only 1/10 (Wilson 95\% CI: [2\%,44\%]), suggesting that systematic reasoning about inverted semantics does not facilitate transfer in the way that code-example mimicry does. The control condition (E32-ctrl: 0/10) confirms that the transfer effect is specific to \emph{inverted} semantics: exposure to non-inverted cross-task code leaves L4 resistance fully intact, ruling out mere cross-task priming as an explanation.

\textbf{Assessment design implication.} L4 resistance depends on \emph{isolation}: if students include inverted code from other tasks in the same prompt (e.g., copying solved examples from a previous assignment), the model can transfer the inversion pattern cross-task. Instructors should ensure that L4 assignments do not provide cross-task inversion examples within the same prompt context. This finding also suggests that L4 resistance is a property of \emph{prompt-level pattern isolation}, not just task-level prior entrenchment.

\subsection{E33: I/O Transparency --- Prior vs.\ Opaque Mapping}
\label{sec:e33_io}

We conducted two experiments to isolate I/O transparency from prior strength. The original E33 design was confounded: the IO-B (\texttt{opaque\_counter}) prompt disclosed explicit algorithmic steps (XOR, shift, conditional add/subtract), making both tasks procedurally transparent; all three models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4o-mini}) achieved 10/10 on both tasks. A replication (E33v2) with a different model set (\texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{o4-mini}) confirmed this result: all models again achieved 10/10 on both IO-A and IO-B, indicating the confound is robust across model generations when algorithmic steps are disclosed. To properly test I/O opacity as an independent variable, we ran E35 (\textbf{redesigned}): IO-B prompts were replaced with pure input/output pairs ($n=8$ examples), with no algorithm disclosed. IO-A remained identical to E33 (explicit ``reverse Fibonacci'' semantics, Fibonacci prior).

\begin{table}[!t]
\centering
\caption{E35 (Redesigned I/O transparency): pass rate on IO-A (strong Fibonacci prior, transparent mapping) vs.\ IO-B (weak prior, opaque I/O only --- no algorithm description). $n=10$ per cell, temperature=0.7.}
\label{tab:e35_io}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Model} & \textbf{IO-A Pass} & \textbf{IO-A Rate} & \textbf{IO-B Pass} & \textbf{IO-B Rate} \\
\midrule
\texttt{gpt-4o}       & 10/10 & \cellcolor{green!15}\textbf{100\%} & 0/10 & \cellcolor{red!10}\textbf{0\%} \\
\texttt{gpt-4.1-mini} & 10/10 & \cellcolor{green!15}\textbf{100\%} & 0/10 & \cellcolor{red!10}\textbf{0\%} \\
\texttt{gpt-4o-mini}  & 5/10  & \cellcolor{yellow!20}50\%           & 0/10 & \cellcolor{red!10}\textbf{0\%} \\
\bottomrule
\end{tabular}%
}
\end{table}

\textbf{Opaque I/O creates universal resistance.} When models receive only input/output pairs with no algorithmic description, all three models fail completely on IO-B (0/30 pooled; Wilson 95\% CI: [0\%,12\%]). By contrast, IO-A---which provides explicit naming cues (``reverse Fibonacci'') that activate a well-known prior---achieves 100\% for both larger models. The IO-A/IO-B contrast (100\% vs.\ 0\% for \texttt{gpt-4o}/\texttt{gpt-4.1-mini}) demonstrates that I/O opacity alone is sufficient to induce resistance, independent of prior strength: the \texttt{opaque\_counter} function has no deep training prior, yet models cannot infer or generalize the underlying bijection from 8 examples to held-out inputs. The \texttt{gpt-4o-mini} IO-A result (5/10) reveals an additional confound: even a transparent I/O mapping fails 50\% when the \emph{Fibonacci prior} interferes with the reversed sequence ordering, demonstrating bidirectional prior effects.

\textbf{Design implication.} L4 resistance is not purely prior-strength-driven; it can also be induced by I/O-opaque task designs for which no training distribution provides a shortcut. This suggests a complementary pathway to NTP-resistant assignment design: rather than requiring the model to violate a known prior (L4), instructors can use opaque I/O specifications that prevent prior activation entirely.

\subsection{E34: Annotation Density Curve (Multi-Model)}
\label{sec:e34_density}

E28 established a non-monotonic annotation density curve for \texttt{o4-mini} alone. E34 extends this to three models across five density levels (0\%, 25\%, 50\%, 75\%, 100\%) on L4 Fibonacci with inverted comparisons ($n=10$ per level per model, 150 total runs). \textbf{Important:} the 0\% level is a \emph{code-example (mimicry) baseline}---it includes an inverted code example rather than I/O-only delivery as in the strict L4 ablation, making it qualitatively distinct from the verbal annotation range (25--100\%); 100\% provides the full explicit rule (L3-equivalent). Results are in Table~\ref{tab:e34_density}.

\begin{table}[!t]
\centering
\caption{E34: Annotation density curve --- pass rate at each density level ($n=10$ per cell, 150 total runs). MAT = Minimum Annotation Threshold (lowest level where pass rate $>$50\%). The 0\% condition includes an inverted \emph{code example} (not I/O-only); it is therefore not comparable to the strict L4 ablation.}
\label{tab:e34_density}
\footnotesize
\begin{tabular}{@{}lccccc|c@{}}
\toprule
\textbf{Model} & \textbf{0\%}$^{\dag}$ & \textbf{25\%} & \textbf{50\%} & \textbf{75\%} & \textbf{100\%} & \textbf{MAT} \\
\midrule
\texttt{gpt-4o}       & 5/10 & 0/10 & 1/10 & 2/10 & 3/10 & None \\
\texttt{gpt-4.1-mini} & \cellcolor{green!15}\textbf{8/10} & 1/10 & 5/10 & 4/10 & 1/10 & 0\%$^{\dag}$ \\
\texttt{o4-mini}      & \cellcolor{green!15}\textbf{7/10} & 3/10 & 6/10 & 6/10 & 1/10 & 0\%$^{\dag}$ \\
\midrule
\multicolumn{7}{l}{\scriptsize $^{\dag}$Code-example (mimicry) baseline; qualitatively distinct from verbal annotation (25--100\%).} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key finding---pattern mimicry via code examples outperforms verbal annotation at all density levels.} All three models show highest pass rates at 0\% annotation density (which includes an inverted \emph{code} example, a qualitatively distinct condition from verbal annotation at 25--100\%), \emph{not} at 100\% (full explicit rule). This is the opposite of the naive expectation that more explicit instructions improve compliance. Within the verbal annotation range (25--100\%), pass rates are non-monotonic: (1)~sharp collapse at 25\% (single comment hint disrupts code-example mimicry without providing sufficient rule content), (2)~partial recovery at 50--75\%, and (3)~collapse at 100\% (full rule triggers prior conflict). The overall pattern distinguishes two qualitatively different mechanisms: code-example mimicry (effective at 0\%) and verbal rule induction (weak and non-monotonic across 25--100\%).

\textbf{Pattern mimicry $>$ rule following.} The 0\% condition succeeds because the inverted code example provides a direct structural template for mimicry---the model copies the comparison operator pattern without needing to understand the underlying rule. Adding explicit annotations paradoxically \emph{hurts} because it shifts the model from pattern mimicry mode to rule-following mode, where the Python prior creates stronger interference. This is consistent with E32's cross-task transfer finding: code examples are more effective than verbal rules for inducing inverted behavior.

\textbf{Contrast with strict L4 ablation.} The 0\% pass rates here (50--80\%) contrast with the strict L4 ablation's 0/350. The critical difference: E34's 0\% prompt shows an inverted \emph{code} example (\texttt{if n >= 0}), while the strict L4 ablation shows only I/O examples and worked solutions without explicitly inverted comparison operators in the prompt code. This confirms that L4 resistance depends on \emph{what} examples show, not just whether a rule is stated.

\textbf{Convergence with E28 (\texttt{o4-mini}).} E34's \texttt{o4-mini} results are directionally consistent with E28: both experiments agree at 50\% annotation density (E28: 6/10; E34: 6/10). The apparent ``peak at 0\%'' in E34 is due to the qualitatively distinct code-example baseline---E28 did not include a 0\% code-example condition, so the two experiments are directly comparable only in the 25--100\% verbal annotation range, where E34 shows the same non-monotonic pattern (50\%/75\% local peak, collapse at 100\%) consistent with E28. Both experiments confirm that annotation saturation hurts reasoning models on prior-entrenched tasks.

\subsection{L4 Hard-Task Extension: Task Complexity vs.\ Prior Entrenchment}
\label{sec:l4_hard}

\begin{table}[!t]
\centering
\caption{L4 hard-task results ($n=10$ per cell, 120 total runs). All tasks use example-only delivery with 2 worked I/O examples. Pass = output matches inverted specification. Models include \texttt{gpt-5.4} and \texttt{o3} (not tested in main L4 ablation). Wilson 95\% CIs at $n=10$: 10/10 $\to$ [72\%,100\%]; 0/10 $\to$ [0\%,28\%]; \texttt{gpt-4o} H1 (4/10) $\to$ [17\%,69\%]---interpret with caution.}
\label{tab:l4_hard}
\footnotesize
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Model} & \textbf{H1: merge\_sort} & \textbf{H2: bin\_search} & \textbf{H3: bfs} & \textbf{Overall} \\
\midrule
\texttt{gpt-4o}   & \cellcolor{yellow!20}4/10  & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & 80\% \\
\texttt{gpt-5.4}  & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \textbf{100\%} \\
\texttt{o3}       & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \textbf{100\%} \\
\texttt{o4-mini}  & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \cellcolor{green!15}10/10 & \textbf{100\%} \\
\bottomrule
\end{tabular}
\end{table}

To test whether more complex algorithms amplify or defeat pattern blindness, we evaluated three hard tasks (H1: descending merge sort, H2: inverted binary search, H3: reversed BFS) across four models including two not previously tested: \texttt{gpt-5.4} and \texttt{o3}. Results are in Table~\ref{tab:l4_hard}.

The results reveal that \textbf{L4 resistance is inversely correlated with task complexity}. Three of four models achieve 100\% pass rate (30/30); \texttt{gpt-4o} achieves 80\% (24/30, with H1 at 4/10). This contrasts sharply with the original L4 results, where \emph{all} models achieve 0\% pass rate on the simple Fibonacci task (T1) under example-only implicit delivery.

\textbf{Methodological note on comparability.} This experiment differs from the strict L4 ablation (\S\ref{sec:l4_ablation}) in two respects, not one: (1)~\emph{prior entrenchment depth}---hard tasks have more varied implementations; and (2)~\emph{I/O specification transparency}---hard-task examples expose the behavioral target unambiguously via input/output pairs (e.g., [3,1,4,2]$\to$[4,3,2,1] makes ``descending order'' self-evident), while the original L4 embeds the inversion as an implicit semantic rule in code structure. These two confounds co-vary: the harder tasks that have shallower priors \emph{also} have more behaviorally transparent I/O specifications. We cannot cleanly separate prior-depth effects from I/O-transparency effects in the current design. Readers should interpret the hard-task results as evidence for \emph{at least one} of these explanatory factors, not necessarily prior depth alone.

\textbf{Prior entrenchment depth as primary explanatory variable (hypothesis).}
We hypothesize that the key variable is \emph{prior entrenchment depth}: how uniformly a single canonical implementation dominates the training distribution. Fibonacci recursion has an extremely deep, uniform Python prior---virtually all training examples use the same \texttt{if n <= 1: return n; return fib(n-1) + fib(n-2)} template. Merge sort, binary search, and BFS admit diverse implementation strategies (iterative vs.\ recursive, different data structures, varied edge-case handling), yielding shallower and more varied priors. When the prior is shallow, the model can override it by extracting the inverted semantics from worked examples; when the prior is deeply entrenched, the NTP distribution overwhelms in-context cues. We label this a hypothesis because the I/O-transparency confound cannot be ruled out with current data.

A complementary observation reinforces this framing: the hard-task prompts provide \emph{explicit I/O specifications} (descending-sorted output, inverted return values, reversed traversal order), while the original L4 tasks embed an \emph{implicit semantic rule} in code structure. Models readily comply with behavioral specifications for complex algorithms, but cannot extract the same inversion when it is embedded implicitly in prior-entrenched code patterns.

\textbf{\texttt{gpt-5.4} and \texttt{o3} hard-task results, and E24 update:} Both achieve perfect 30/30 on the hard-task battery (shallow-prior, I/O-transparent tasks). To directly test whether these models also overcome \emph{deep} prior entrenchment, we ran E24: strict example-only Variant~A on L4 Fibonacci for both models ($n=10$ each). Result: \texttt{gpt-5.4} 0/10 (PPR=1.00) and \texttt{o3} 0/10 (PPR=1.00). Frontier models are defeated by the same Fibonacci task that defeats all seven core models. The open question from the hard-task section is now resolved: \emph{pattern blindness on prior-entrenched tasks holds for current frontier models} as of 2026-03-25.
