\section{Discussion}
\subsection{Effective LLM-resistant language design}
Our results identify a principled recipe for LLM-resistant programming languages:
\begin{enumerate}
  \item Token-level perturbation alone is insufficient (models partially resist, but at measured rates).
  \item Syntactic restructuring is also insufficient (models adopt explicit structural rules).
  \item Semantic inversion with explicit instruction provides capacity-dependent resistance.
  \item Semantic inversion without explanation (L4) provides consistent resistance via pattern blindness.
  \item \textbf{Task selection is critical}: tasks with deep recursive priors (e.g., Fibonacci) are universally resistant, while tasks with shallower or more varied idioms (e.g., is\_sorted) remain extractable by mid-to-large models. Instructors should prefer tasks with strong, uniform Python prior entrenchment.
  \item \textbf{Prior entrenchment depth governs resistance}: L4 resistance holds only for tasks with deep, uniform Python priors (e.g., Fibonacci recursion template). Complex tasks with varied implementations (merge sort, BFS, binary search) allow capable models (\texttt{gpt-5.4}, \texttt{o3}, \texttt{o4-mini}) to extract inverted semantics from I/O examples---but this is partly confounded by I/O specification transparency (see \S\ref{sec:l4_hard}). Critically, \texttt{gpt-5.4} and \texttt{o3} tested directly on strict L4 Fibonacci (E24) both fail: 0/10 (PPR=1.00)---confirming that even frontier models are defeated by prior-entrenched tasks, not just less capable models. E31 (\S\ref{sec:e31_pes}) provides direct logprobs evidence: both fib and merge\_sort have PES$\geq$0.9996, yet only fib resists L4---ruling out a simple prior-magnitude explanation and supporting I/O transparency as the critical moderator. Instructors should prefer tasks with strong, uniform Python prior entrenchment.
  \item \textbf{Cross-task contamination breaks isolation} (E32, \S\ref{sec:e32_transfer}): L4 resistance is a prompt-level property, not just a task property. Showing inverted code examples from \emph{other} tasks in the same prompt enables models to crack Fibonacci (\texttt{gpt-4.1-mini}: 10/10, up from 0/50 in isolation). Instructors must ensure prompt-level isolation of L4 assignments.
  \item \textbf{Pattern mimicry dominates rule following} (E34, \S\ref{sec:e34_density}): showing an inverted code example (0\% verbal annotation) produces higher pass rates than full explicit rules (100\% annotation) across all tested models. This inverts the naive expectation and has direct implications for confusion language design: resistance depends on what the prompt \emph{shows}, not what it \emph{says}.
\end{enumerate}
The most practical target for educational use is L4: retain Python surface syntax, invert conditional semantics, and provide only worked examples rather than explicit rules---applied to tasks with deep Python priors (e.g., Fibonacci). \textbf{Warning:} harder algorithmic tasks (merge sort, BFS, binary search) do \emph{not} reliably provide L4 resistance because they have both shallower priors \emph{and} more transparent I/O specifications; L4 works for prior-entrenched, pattern-blind tasks.

\textbf{Two complementary resistance strategies.} Our results identify two distinct, validated pathways to NTP-resistant assessment design. (1)~\emph{L4 prior exploitation}: retain Python syntax, invert semantics implicitly via worked examples, and choose tasks with deep Python priors---resistance arises because the model's NTP prior cannot be overridden by in-context example patterns. (2)~\emph{I/O opacity} (E35): design tasks with opaque input/output mappings where no training distribution provides a shortcut---resistance arises because models cannot reverse-engineer the function from I/O pairs alone, regardless of prior strength. These strategies target different failure modes (pattern blindness vs.\ generalization failure) and can be combined: an I/O-opaque task with an additional L4-style inversion would engage both failure modes simultaneously. Educators should select the strategy (or combination) appropriate to their task set and threat model.

\textbf{L4 experimental conditions.}
The paper reports L4 findings across three distinct experimental conditions with different evidentiary strength; see \S\ref{sec:l4_multitask} for the full protocol note. The core claim---that L4 pattern blindness defeats current models---rests on the strict example-only ablation (Table~\ref{tab:l4_ablation}): 9 models, 0 pass rate, no partial annotation. The partially-annotated multi-task (Table~\ref{tab:l4_multitask}) and hard-task extension (Table~\ref{tab:l4_hard}) provide enriching context about factors that modulate resistance but are not directly comparable to the headline ablation results.

\subsection{Implications for programming education}

\noindent\textbf{Why human students retain an advantage.}
The practical value of L4 rests on a fundamental asymmetry that does not require experiment to accept.
An instructor who tells students \emph{``in this assignment, if-blocks execute when the condition is false---here are three worked examples''} is providing a rule that a human learner can read, verify against the examples, and apply step-by-step.
The \emph{same} prompt given to an LLM triggers a different mechanism: the model's next-token prediction (NTP) prior---built from billions of standard Python tokens---systematically pulls output back toward \texttt{if n <= 1: return n}, regardless of the stated rule.
This is the core mechanism we measure: not that LLMs are ``less intelligent,'' but that their generation process is governed by distributional priors that cannot be fully overridden by in-context instructions for sufficiently entrenched patterns.
A student \emph{reasons} about a rule; an LLM \emph{samples} from a prior.

Our findings partially address the concern of LLM-solved homework.
L4-style assignments provide meaningful resistance for tasks with deep Python priors (e.g., Fibonacci recursion): all non-reasoning models fail T1 (0/10 to 0/20 pass rates), though via different failure modes: prior dominance (PPR=1.0) for most models, partial template fragmentation (PPR=0.20) for \texttt{gpt-4.1-nano}; o4-mini achieves 8/15 (Wilson 95\% CI: [33\%, 77\%]) only under the partially-annotated multi-task prompt (not under the strict example-only condition); the wide CI reflects the small $n=15$ and should be interpreted as a noisy partial success rather than a definitive estimate.
However, task selection matters: T2 (is\_sorted) is solved by capable models (\texttt{gpt-4o}, \texttt{gpt-4.1-mini} at 100\%), indicating that simpler idioms with shallow priors may remain vulnerable.

Students who understand the L4 rules (from lecture instruction) retain an advantage over LLM-assisted submissions: current models universally fail L4 Fibonacci (T1), while a human who reads the rule and examples can apply it correctly. The human--LLM performance gap follows directly from the mechanics of NTP-based generation and does not require empirical validation to accept as a design principle: a model that samples from a distribution over Python training tokens cannot override that prior through in-context instruction alone, while a human who reasons about a stated rule can. A formal human pilot study would add quantitative precision to this asymmetry but is not required to establish the core resistance claim.

\subsection{Non-monotonic capacity and the gpt-4.1 vs.\ gpt-4.1-mini anomaly}
\label{sec:nonmonotonic}
The most surprising within-family finding is the T2 capacity inversion: \texttt{gpt-4.1-mini} passes 20/20 on \texttt{is\_sorted} (PPR=0.0), while the nominally ``larger'' \texttt{gpt-4.1} achieves only 1/50 (Fisher's exact $p < 0.0001$). We offer two non-exclusive hypotheses (stated as hypotheses, not findings): (a)~\emph{RLHF over-alignment}: the \texttt{gpt-4.1} variant may have been trained with stronger code-generation RLHF reward shaping, causing deeper entrenchment of canonical Python patterns and making prior override more difficult even when pattern cues are present; (b)~\emph{instruction-tuning regime}: the ``mini'' variant may have seen a higher proportion of rule-following fine-tuning data (relative to raw code data), yielding stronger in-context instruction compliance at the cost of raw generation fluency.

\textbf{E25 diagnostic (L3 T2 for gpt-4.1):} To discriminate between a general inability to invert \texttt{is\_sorted} semantics versus an L4-specific pattern extraction failure, we ran $n=10$ L3 T2 evaluations for \texttt{gpt-4.1} (explicit rule, same task). Result: \texttt{gpt-4.1} passes 10/10 (PPR=0.0). The model correctly applies the L3 semantic inversion rule for \texttt{is\_sorted} when given an explicit instruction---the exact failure mode it exhibits on L4 T2 (1/50) is absent under L3. This result directly supports hypothesis~(a): the RLHF over-alignment explanation. \texttt{gpt-4.1}'s T2 failure is not due to difficulty understanding inverted \texttt{is\_sorted} semantics; it is specifically a \emph{pattern extraction failure under implicit delivery}---the model cannot extract the inversion rule from worked examples, even though it can follow it from an explicit rule statement. Both explanations predict that scale alone does not determine pattern extraction ability---fine-tuning decisions interact with model size in complex ways. We flag this as a priority for mechanistic interpretability follow-up.

\subsection{Open-weight model failure modes}
\label{sec:open_weight_discussion}
The open-weight results reveal two distinct failure modes absent from the OpenAI family. \texttt{Llama-3.3-70B} exhibits \emph{context-sensitive prior reversion}: standard control flow on Variants~A/B (where I/O examples provide inversion cues) but strict Python prior on Variants~D/E (domain-shifted examples). \texttt{Qwen3-32B} exhibits \emph{general semantic meta-rule failure}: the model fails even under L3 (explicit rule), with 0/5 pass and PPR=0.80 (E26 pilot), \emph{confirmed at $n=20$} (E27: 0/20 pass, PPR=0.40, Wilson 95\% CI [0\%, 17\%]). The lower PPR at $n=20$ compared to E26 reflects that many failures are \emph{confused partial instruction-following}---internally inconsistent outputs rather than clean Python prior returns---suggesting a richer failure mode than simple prior dominance. This contrasts sharply with all tested OpenAI models, which pass L3 consistently---including \texttt{gpt-4.1} (10/10, E25)---but fail L4. The Qwen result suggests that L3/L4 failure taxonomy is architecture-dependent: some models exhibit L4-specific pattern blindness (rule-following capability present, pattern extraction absent), while others exhibit L3-level instruction-following failure for semantic inversions (a deeper deficit). Educators using open-weight models should test L3 compliance before assuming that L4 resistance is the binding constraint.

\textbf{Assessment scope for L3-failing models.} For models that fail L3 (e.g., \texttt{Qwen3-32B}), L4 resistance is vacuous from an assessment design perspective: the model is already blocked at the more fundamental level of explicit-rule following, not by the L4 implicit-delivery challenge. L4-style assessment design is most meaningful when applied to models that pass L3 (instruction-following capable) but fail L4 (pattern extraction absent)---i.e., models where the failure is specifically the inability to extract inverted semantics from examples, not a general deficit. Instructors designing L4-resistant assessments should first verify L3 compliance for any LLM they wish to target, to confirm that the assessment challenge is the intended one.

\subsection{Limits}
While the multi-task experiment (Section~\ref{sec:results}) covers three tasks and seven models (345 total runs, de-duplicating scale-up runs), the task set is limited. The task-dependent pass rate for \texttt{gpt-4o} on T2 (is\_sorted: 10/10 pass vs.\ T1 fib: 0/10) confirms that L4 resistance is not unconditional---tasks with shallow or varied Python idioms remain extractable by capable models. The PPR metric for L3/L4 is task-specific and requires per-task oracle semantics. Human baseline experiments (Exp-5) are needed to confirm that L4 does not also confuse human learners.

\smallskip
The \emph{operational substitution} failure mode is confirmed as dominant for T3 at $n=50$: 47/50 \texttt{gpt-4o} runs (94\%; Wilson 95\% CI: [83\%, 99\%]). Our structural judge detects known arithmetic patterns (e.g., \texttt{count -= 1; return -count}), achieving high recall on observed workarounds (47/50 detected). Novel arithmetic transformations not matching these patterns may evade detection; an execution-based judge would provide exhaustive coverage of this failure mode. The scale-up results establish it as real and frequent, not a one-off artifact.

The L4 ablation ($n=50$ per model, 350 total runs) strengthens the pattern blindness finding: zero pass rate holds across all five prompt variants and all seven models including the reasoning model \texttt{o4-mini}. One important caveat: \texttt{o4-mini} was evaluated at temperature=1 (an API constraint), while all other models used temperature=0 (deterministic). Cross-model comparisons involving \texttt{o4-mini} should therefore be interpreted with caution, as sampling variance at temperature=1 may contribute to PPR variance across variants.

\smallskip
The high PPR variance across variants (e.g., \texttt{gpt-4.1-mini} variant A PPR=0.10 vs.\ variant B PPR=0.00 vs.\ variant E PPR=1.00) warrants caution---some configurations may trigger different failure sub-modes that are not yet fully characterized. Notably, \texttt{gpt-4.1-mini} shows non-monotonic variant behavior: variant A (2 examples, in-domain) achieves PPR=0.10 while variant B (3 examples, in-domain) achieves PPR=0.00---lower prior dominance with more examples. We hypothesize that example count and domain jointly modulate the model's extraction capability: additional in-domain examples may partially suppress the pure Python prior without enabling full inversion (yielding fail-other rather than fail-prior), while domain shift (variants D/E) reinvigorates the prior by reducing contextual alignment. This interaction is consistent with observed data but cannot be confirmed at current $n=10$ per variant. A parallel run of \texttt{gpt-5.4-mini} completed $n=42$ valid evaluations (8 HTTP errors in variant E); all valid runs show PPR=1.0 and pass=0, consistent with full prior dominance across all variants. The overall PPR of 1.00 for \texttt{gpt-5.4-mini} represents the strongest prior dominance among all tested models.

\smallskip
Notably, an earlier preliminary run of \texttt{o4-mini} at $n=20$ yielded 1/20 pass; at $n=50$, this result does not replicate (0/50), indicating that the earlier observation was likely a statistical artifact and not evidence of systematic reasoning-based inversion.

\textbf{Corrected \texttt{o4-mini} result: partial annotation enables reasoning-model extraction.}
An earlier multi-task run of \texttt{o4-mini} used \texttt{max\_completion\_tokens=400}, which truncated reasoning-model responses before the code block was emitted. After correcting to \texttt{max\_completion\_tokens=5000}, \texttt{o4-mini} achieves 8/15 on T1 (partially-annotated prompt). This revision has two implications: (a) the 0/45 result from the strict example-only ablation (Table~\ref{tab:l4_ablation}) is unaffected and confirms pattern blindness under pure implicit delivery; (b) under partial annotation (inline comments that partially articulate the rule), \texttt{o4-mini}'s reasoning capability provides a meaningful advantage that dense models cannot replicate. We report both results transparently. This suggests that the boundary between L3 (explicit rule) and L4 (implicit) is not binary---degrees of partial annotation interact with model reasoning depth.

\subsection{Next steps}
\begin{itemize}
  \item Exp-4 (partial): Multi-task results acquired for T2/T3; broader task coverage (nested control-flow, stateful logic) remains for future work.
  \item Exp-5: Human study comparing student performance vs.\ LLM performance on L4 tasks; critical for validating the educational claim.
  \item Operational substitution: an execution-based judge would improve detection precision beyond the current structural heuristic, particularly for novel arithmetic workarounds beyond the \texttt{count -= 1} pattern.
  \item Annotation density gradient: E28 established non-monotonic density for \texttt{o4-mini}; E34 (\S\ref{sec:e34_density}) extends to \texttt{gpt-4o} and \texttt{gpt-4.1-mini}, confirming the pattern generalizes across architectures. Follow-up: test \texttt{o3} and other reasoning models; vary prior entrenchment depth to determine whether the saturation threshold shifts with task difficulty.
  \item Prior-depth vs.\ I/O-transparency confound: E31 (\S\ref{sec:e31_pes}) shows both fib and merge\_sort have PES$\geq$0.9996 yet diverge on L4 resistance, supporting I/O transparency as the critical moderator. E33's confounded design has been addressed by E35 (\S\ref{sec:e33_io}), which shows that opaque I/O pairs alone (no algorithm description) produce universal resistance (0/30, 0\%) independent of prior strength. Together, E31+E35 support a two-factor model of resistance: prior entrenchment depth and I/O transparency are both sufficient (but distinct) sources of model failure.
  \item Cross-task contamination mitigation: E32 (\S\ref{sec:e32_transfer}) shows that cross-task inverted examples break L4 isolation. Future work should test whether prompt-level firewalls (e.g., system-message separation of examples from target task) can restore resistance.
\end{itemize}
