\section{Prompt and Evaluation Protocol (Complete)}

\subsection{L1: Token Alias Substitution}

\subsubsection{System prompt (baseline S0)}
\begin{quote}
\small
Solve the given prompt exactly. Return ONLY code (no explanation, no markdown). Follow alias mapping strictly.
\end{quote}

\subsubsection{System prompt (context-pack S1)}
\begin{quote}
\small
Solve the given prompt exactly. Return ONLY code (no explanation, no markdown). Follow alias mapping strictly.

[Transformed language contract]\\
Use transformed aliases exactly as follows: (token$\rightarrow$alias list from prompt).\\
Do NOT output original Python keywords for transformed entries.\\
Banned originals for this task: (derived from mapping).\\
Output only executable code in the transformed language.
\end{quote}

\subsubsection{User prompt structure (verbatim template)}
\begin{quote}
\small
\textit{[Language Description]}\\
Python grammar remains unchanged.\\
Keyword aliases are user-configurable strings (including phrases with spaces).
At runtime, detected aliases are normalized back to Python via token replacement.\\
\\
\textit{Current mapping summary:}\\
\quad- if $\rightarrow$ \{alias\}\\
\quad- in $\rightarrow$ \{alias\}\\
\quad- def $\rightarrow$ \{alias\}\\
\quad- for $\rightarrow$ \{alias\}\\
\quad- return $\rightarrow$ \{alias\}\\
\quad- print $\rightarrow$ \{alias\}\\
\\
\textit{[Task]}\\
1) Define function \texttt{fib(n)}.\\
2) If \texttt{n <= 1}, return \texttt{n}.\\
3) Otherwise return \texttt{fib(n-1) + fib(n-2)}.\\
4) Loop \texttt{i = 0..6}, print \texttt{(i, fib(i))}.
\end{quote}

\subsubsection{Representative alias mappings across prompt versions}
Three representative prompt instances from the L1 bank:

\textbf{v1}: if$\to$\texttt{return}, def$\to$\texttt{if}, for$\to$(long phrase), in$\to$(long phrase), return$\to$(long phrase), print$\to$\texttt{do}

\textbf{v35}: in$\to$(phrase), def$\to$\texttt{def}, for$\to$\texttt{xk}, return$\to$(long phrase), if$\to$(anti-prior phrase)

\textbf{v90}: def$\to$\texttt{return}, for$\to$(phrase), in$\to$(phrase), return$\to$\texttt{out}, if$\to$(long phrase), print$\to$\texttt{qn}

\subsection{L2: Syntax Inversion}

\subsubsection{Grammar specification (given to model)}
\begin{quote}
\small
This is NOT Python. It uses inverted syntax.\\
ALL standard Python keywords (def, return, if, for, in) are FORBIDDEN.

Grammar rules:\\
\quad- Function definition: \texttt{:define name [args] ->} (not \texttt{def name(args):})\\
\quad- Return: \texttt{<= value} (not \texttt{return value})\\
\quad- Conditional: \texttt{value ?cond: result ;else: alt ;end}\\
\quad- Loop: \texttt{loop var over iterable \{ body \}}\\
\quad- Block end: \texttt{;;}
\end{quote}

\subsubsection{Task (same across L1--L4)}
Same Fibonacci function (fib(0)=0, fib(1)=1, fib(n)=fib(n-1)+fib(n-2), print i=0..6).

\subsection{L3: Semantic Inversion (Explicit Rule)}

\subsubsection{System prompt}
\begin{quote}
\small
Follow the L3 language rules exactly. If-blocks run when condition is FALSE.
\end{quote}

\subsubsection{User prompt (complete)}
\begin{quote}
\small
\textit{L3 is a Python-LIKE language with semantic inversion of conditionals.}\\
The syntax is identical to Python, but:\\
RULE: The indented block of an if-statement executes when the condition is FALSE.\\
RULE: The else-block executes when the condition is TRUE.\\
RULE: This is opposite to Python's behavior.\\
\\
Example:\\
\texttt{if x > 5:}\\
\quad\texttt{print("small")} \textit{(runs when x is NOT > 5)}\\
\texttt{else:}\\
\quad\texttt{print("large")} \textit{(runs when x IS > 5)}\\
\\
Task: Write a Fibonacci function in L3. The intended logic is:\\
if n <= 1, return n; else return fib(n-1)+fib(n-2).\\
Remember: your if-block runs when condition is FALSE.\\
Loop i=0..6, print (i, fib(i)).
\end{quote}

\subsection{L4: Semantic Inversion (Implicit, Example-Only)}

\subsubsection{System prompt}
None (user-role only).

\subsubsection{User prompt (complete)}
\begin{quote}
\small
Here are some examples of code in our programming language:\\
\\
Example 1 -- Simple max function:\\
\texttt{def max\_val(a, b):}\\
\quad\texttt{if a > b:}\\
\quad\quad\texttt{return b}\ \textit{(returns smaller value when a > b)}\\
\quad\texttt{return a}\\
Output: max\_val(3, 5) = 3, max\_val(7, 2) = 2\\
\\
Example 2 -- Even check:\\
\texttt{def is\_even(n):}\\
\quad\texttt{if n \% 2 == 0:}\\
\quad\quad\texttt{return False}\\
\quad\texttt{return True}\\
Output: is\_even(4) = False, is\_even(3) = True\\
\\
Now write the Fibonacci function in this same language:\\
- fib(0)=0, fib(1)=1, fib(n)=fib(n-1)+fib(n-2)\\
- Print fib(0) through fib(6)
\end{quote}

\subsection{Judge Protocol (L1)}

\begin{enumerate}
  \item \textbf{Alias compliance}: for each required token (\texttt{if}, \texttt{in}, \texttt{def}, \texttt{for}, \texttt{return}), check if alias is absent from output.
  \item \textbf{Keyword leakage}: check if original token appears using word-boundary regex.
  \item \textbf{Normalization}: replace alias with original token via regex/string replacement.
  \item \textbf{Parse check}: run \texttt{ast.parse()} on normalized code.
  \item \textbf{Skeleton check}: verify fib function, base case, recursion, and print loop.
\end{enumerate}

Metrics:
\begin{itemize}
  \item KLR = (leaked tokens) / (active alias slots)
  \item ACR = 1 $-$ KLR
  \item PSS = 0.4$\times$ACR + 0.25$\times$parse\_ok + 0.25$\times$skeleton\_score + 0.1$\times$loop\_ok
  \item PIR = KLR / (KLR + ACR)
\end{itemize}

\subsection{Judge Protocol (L3/L4)}

Binary check:
\begin{itemize}
  \item \textbf{Inverted (correct)}: model writes \texttt{if n > 1:} or \texttt{if n >= 2:} --- inverted from Python-expected \texttt{if n <= 1:}
  \item \textbf{Python prior (incorrect)}: model writes \texttt{if n <= 1: return n} without inversion
  \item PPR = fraction of samples using Python-prior pattern
\end{itemize}

\subsection{Execution Commands (Complete)}

\subsubsection{L1 full-range baseline}
\begin{verbatim}
./scripts/use_avrtg_openai_key.sh env \
  CONTEXT_PACK=0 EVAL_MODEL=gpt-4o-mini \
  RESULT_JSON=docs/research/results/\
  prompt-batch-v1-120.gpt4o-mini.2026-03-23.json \
  python3 scripts/run_gpt54_eval.py
\end{verbatim}

\subsubsection{L1 context-pack ablation}
\begin{verbatim}
CONTEXT_PACK=1 EVAL_MODEL=gpt-4o-mini ...
\end{verbatim}

\subsubsection{L3 experiment}
\begin{verbatim}
./scripts/use_avrtg_openai_key.sh env \
  EVAL_MODEL=gpt-4o python3 -c "...l3_eval..."
\end{verbatim}

\subsubsection{Partial judge (re-evaluate any batch)}
\begin{verbatim}
python3 scripts/partial_judge.py \
  docs/research/results/prompt-batch-v1-120.gpt4o-mini.json
\end{verbatim}

\subsubsection{Run comparison}
\begin{verbatim}
python3 scripts/compare_eval_runs.py \
  --base ..baseline.json --cand ..contextpack.json \
  --out-json ..compare.json --out-md ..compare.md
\end{verbatim}
