\section{Complete Reproducibility Protocol}
\subsection{Prompt contract (fully written)}
All runs use the following prompt contract.

\textbf{System prompt (baseline):}
\begin{quote}
\small Solve the given prompt exactly. Return ONLY code (no explanation, no markdown). Follow alias mapping strictly.
\end{quote}

\textbf{System prompt (context-pack extension):}
\begin{quote}
\small
[Transformed language contract]
Use transformed aliases exactly as follows: (token\(\rightarrow\)alias list from the prompt).
Do NOT output original Python keywords for transformed entries.
Output only executable code in the transformed language.
\end{quote}

\textbf{User prompt template:}
\begin{quote}
\small
\textit{Language Description}: Python grammar remains unchanged.

\textit{Current mapping summary}:
\begin{itemize}
  \item if \(\rightarrow\) \{alias\}
  \item in \(\rightarrow\) \{alias\}
  \item def \(\rightarrow\) \{alias\}
  \item for \(\rightarrow\) \{alias\}
  \item return \(\rightarrow\) \{alias\}
  \item print \(\rightarrow\) \{alias\}
\end{itemize}

\textit{Task}:
\begin{enumerate}
  \item Define function \texttt{fib(n)}.
  \item If \texttt{n <= 1}, return \texttt{n}.
  \item Otherwise return \texttt{fib(n-1) + fib(n-2)}.
  \item Loop \texttt{i=0..6} and print \texttt{(i, fib(i))}.
\end{enumerate}
\end{quote}

\subsection{Judge details (deterministic)}
For each sample, the judge applies:
\begin{enumerate}
  \item alias compliance and original keyword leakage checks,
  \item alias\(\rightarrow\)Python normalization and AST parseability,
  \item benchmark skeleton checks (\texttt{fib}, base case, recursion, loop-print contract).
\end{enumerate}
Pass is defined as zero violations.

\subsection{Exact run settings reported in this paper}
Representative full-range runs:
\begin{itemize}
  \item \texttt{prompt-batch-v1-120.gpt4o-mini.2026-03-23.json}
  \item \texttt{prompt-batch-v1-120.gpt41-mini.2026-03-23.json}
  \item \texttt{prompt-batch-v1-120.gpt41-nano.2026-03-23.json}
  \item \texttt{prompt-batch-v1-120.gpt54-mini.2026-03-23.json}
\end{itemize}
Context-pack ablation:
\begin{itemize}
  \item baseline: \texttt{prompt-batch-v1-120.gpt4o-mini.2026-03-23.json}
  \item context-pack: \texttt{prompt-batch-v1-120-contextpack.gpt4o-mini.2026-03-23.json}
  \item diff report: \texttt{prompt-batch-compare-contextpack-vs-base.gpt4o-mini.2026-03-23.*}
\end{itemize}

\subsection{Execution settings (compact)}
To avoid line-break issues, we report settings in compact form:
\begin{itemize}
  \item model: \texttt{gpt-4o-mini}
  \item prompt range: \texttt{v1--v120}
  \item context-pack: \texttt{on/off} ablation
  \item failure cap: \texttt{999}
  \item outputs: run JSON + comparison JSON/MD
\end{itemize}
Full runnable commands are provided in repository docs/research artifacts.

\subsection{What can be reproduced from this paper alone}
A reader can reproduce:
\begin{itemize}
  \item full prompt-batch runs (v1--v120),
  \item strict pass/fail and violation logs,
  \item context-pack ablation deltas,
  \item artifact-level evidence mapping from claims to files.
\end{itemize}
