\section{Appendix: Prompt Samples, Taxonomy, and Config Snapshot}
\subsection{Prompt samples (readable form)}
Representative prompt-map blocks from the versioned bank:

\textbf{Sample A (v1)}
\begin{itemize}
  \item if \(\rightarrow\) return
  \item def \(\rightarrow\) if
  \item for \(\rightarrow\) long natural phrase (renamed symbol)
  \item in \(\rightarrow\) long natural phrase (strict compilation token)
  \item return \(\rightarrow\) long natural phrase (keyword mapping token)
  \item print \(\rightarrow\) do
\end{itemize}

\textbf{Sample B (v35)}
\begin{itemize}
  \item in \(\rightarrow\) mapped token phrase
  \item def \(\rightarrow\) def
  \item for \(\rightarrow\) xk
  \item return \(\rightarrow\) long phrase preserving grammar
  \item if \(\rightarrow\) anti-prior phrase
\end{itemize}

\textbf{Sample C (v90)}
\begin{itemize}
  \item def \(\rightarrow\) return
  \item for \(\rightarrow\) phrase mapped back to reserved token
  \item in \(\rightarrow\) exact mapping phrase
  \item return \(\rightarrow\) out
  \item if \(\rightarrow\) long natural phrase
  \item print \(\rightarrow\) qn
\end{itemize}

\subsection{Violation taxonomy summary (gpt-4o-mini, v1--v120)}
From the context-pack comparison artifact, dominant violation categories are:
\begin{itemize}
  \item Original keyword leakage (e.g., \texttt{for}, \texttt{if}, \texttt{return}, \texttt{in}, \texttt{def})
  \item Parse failure after normalization (line-specific syntax failure)
  \item Missing alias emissions for required transformed tokens
\end{itemize}

Top counts in context-pack run include:
\begin{itemize}
  \item \texttt{for} leaked: 52
  \item parse failed (line 1): 35
  \item \texttt{if} leaked: 33
\end{itemize}

\subsection{PSS (Partial Structural Score) Supplementary Data}
\label{sec:appendix_pss}
PSS is a heuristic composite metric (alias compliance 0.4, parse 0.25, skeleton 0.25, format 0.1) reported as a secondary diagnostic only. Weights are not validated; PSS is not used as a primary result. The table below provides reference values alongside KLR for L1 evaluation.

\begin{table}[h]
\centering
\caption{L1 KLR and PSS (supplementary). PSS is exploratory; KLR is the validated primary metric.}
\label{tab:l1_pss}
\footnotesize
\begin{tabular}{@{}llcc@{}}
\toprule
\textbf{Model} & \textbf{Delivery} & \textbf{KLR}$\downarrow$ & \textbf{PSS}$\uparrow$ \\
\midrule
\texttt{gpt-4.1-mini} & ctx-pack & 0.21 & 0.37 \\
\texttt{Llama-3.3-70B}$^\star$ & ctx-pack & 0.23 & 0.34 \\
\texttt{gpt-4o-mini}  & ctx-pack & 0.26 & 0.32 \\
\texttt{o4-mini}      & ctx-pack & 0.27 & 0.41 \\
\texttt{gpt-4.1}      & ctx-pack & 0.30 & \textbf{0.43} \\
\texttt{gpt-4o}       & baseline & 0.37 & 0.40 \\
\texttt{gpt-4.1-nano} & ctx-pack & 0.38 & 0.30 \\
\texttt{gpt-4o-mini}  & baseline & 0.42 & 0.27 \\
\texttt{gpt-4.1-mini} & baseline & 0.43 & 0.33 \\
\texttt{gpt-5.4-mini} & baseline & 0.67 & 0.25 \\
\texttt{gpt-4.1-nano} & baseline & 0.69 & 0.16 \\
\texttt{Qwen3-32B}$^\star$ & ctx-pack & 1.00 & 0.00 \\
\bottomrule
\multicolumn{4}{l}{\scriptsize $^\star$Open-weight via Groq API; $n=20$, ctx-pack only.} \\
\end{tabular}
\end{table}

\subsection{L4 Multi-Task Full PPR Values}
\label{sec:appendix_l4_multitask_ppr}
Full Python Prior Rate (PPR) values for Table~\ref{tab:l4_multitask}. PPR = fraction of outputs matching canonical Python prior (lower = more L4-compliant). Pass rate = 0 for all cells except \texttt{gpt-4o} T2, \texttt{gpt-4.1-mini} T2, and \texttt{o4-mini} T1.

\begin{table}[h]
\centering
\caption{L4 multi-task full PPR. $\dagger$~Partially-annotated prompts (not comparable to Table~\ref{tab:l4_ablation}).}
\label{tab:appendix_l4_ppr}
\footnotesize
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}l ccc ccc ccc@{}}
\toprule
& \multicolumn{3}{c}{\textbf{T1: fibonacci}} & \multicolumn{3}{c}{\textbf{T2: is\_sorted}} & \multicolumn{3}{c}{\textbf{T3: count\_vowels}} \\
\cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
\textbf{Model} & $n$ & \textbf{Pass} & \textbf{PPR} & $n$ & \textbf{Pass} & \textbf{PPR} & $n$ & \textbf{Pass} & \textbf{PPR} \\
\midrule
\texttt{gpt-4o}       & 10 & 0/10  & 1.00 & 10 & 10/10 & 0.00 & 50 & 0/50 & 0.00 \\
\texttt{gpt-4o-mini}  & 10 & 0/10  & 0.00 & 10 & 0/10  & 1.00 & 10 & 0/10 & 0.00 \\
\texttt{gpt-4.1}      & 10 & 0/10  & 1.00 & 50 & 1/50  & 0.42 & 10 & 0/10 & 0.00 \\
\texttt{gpt-4.1-mini} & 20 & 0/20  & 1.00 & 20 & 20/20 & 0.00 & 20 & 0/20 & 0.25 \\
\texttt{gpt-4.1-nano} & 10 & 0/10  & 0.20 & 10 & 0/10  & 1.00 & 10 & 0/10 & 0.00 \\
\texttt{gpt-5.4-mini} & 10 & 0/10  & 1.00 & 10 & 0/10  & 1.00 & 10 & 0/10 & 0.00 \\
\texttt{o4-mini}      & 15 & 8/15  & 0.33 & 20 & 0/20  & 0.30 & 20 & 0/20 & 0.00 \\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Execution config snapshot (non-secret)}
\begin{table}[h]
\caption{Key runtime settings}
\label{tab:config_snapshot}
\footnotesize
\centering
\begin{tabular}{@{}lp{3.5cm}@{}}
\toprule
Setting & Value \\
\midrule
Models & gpt-4o, gpt-4o-mini, gpt-4.1, gpt-4.1-mini, gpt-4.1-nano, gpt-5.4-mini, o4-mini \\
L1 range & v1--v120 (baseline); v1--v20 (ctx-pack for 4.1/o4-mini) \\
CONTEXT\_PACK & 0/1 \\
Temperature & 0 (all models except o4-mini); 1 (o4-mini, API requirement) \\
Judge & deterministic \\
\bottomrule
\end{tabular}
\end{table}

Secret-bearing values (API keys) are excluded and loaded via runtime environment.
Language specs are in \texttt{language-levels/} (repository).

\subsection{Experiment Log: E31--E34 (2026-03-25)}
\label{sec:appendix_e31_e34}

\textbf{E31 (Prior Entrenchment Score).} Logprobs API query at base-case decision point (\texttt{max\_tokens=1}, \texttt{top\_logprobs=5}, \texttt{temperature=0}). Models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4o-mini}. Tasks: fib, merge\_sort, is\_sorted, count\_vowels. 12 API calls total. \texttt{gpt-4o-mini} excluded from analysis (emits markdown backtick tokens rather than code tokens at this position). Results: \texttt{docs/research/results/e31-prior-entrenchment-score.json}.

\textbf{E32 (Cross-Task Transfer).} Few-shot prompts with inverted merge\_sort and binary\_search code, then fib request. Models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{o4-mini}. $n=10$ per model, temperature=0.7. Judge: regex check for inverted comparison operators ($>$, $\geq$) in base case. Initial \texttt{o4-mini} run yielded HTTP 400 (missing \texttt{max\_completion\_tokens}); re-run with corrected parameters yielded 1/10. Results: \texttt{docs/research/results/e32-cross-task-transfer.json}.

\textbf{E33 (I/O Transparency).} Two tasks: IO-A (reverse\_fib, transparent) and IO-B (opaque\_counter, opaque). Models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{gpt-4o-mini}. $n=10$ per task per model, temperature=0.7. Null result: 10/10 on both tasks for all models. Design flaw: IO-B prompt included explicit algorithm, making it procedurally transparent. Results: \texttt{docs/research/results/e33-io-transparency.json}.

\textbf{E34 (Annotation Density Curve).} Five density levels (0\%, 25\%, 50\%, 75\%, 100\%) on L4 Fibonacci with inverted comparisons. Models: \texttt{gpt-4o}, \texttt{gpt-4.1-mini}, \texttt{o4-mini}. $n=10$ per level per model (150 total). Temperature=0.7. Judge: regex check for inverted comparison in base case. Results: \texttt{docs/research/results/e34-annotation-density.json}.
